### 2022-02-05 Codes/PrepareResults.Rmd
initial submission

---
title: "Machine Learning Methods for Site-specific Input Management"
author: Shunkei Kakimto
output: word_document
---

```{r}
library(knitr)
# library(here)

knitr::opts_chunk$set(
    cache = FALSE,
    warning = FALSE,
    message = FALSE
    )

# opts_knit$set(root.dir=here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/ML_VRA")
```

```{r, cache = FALSE}
# === packages ===#
library(sf)
library(data.table)
library(RColorBrewer)
library(patchwork)
library(grf)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(viridis)
library(tidyverse)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(here)
library(gridExtra)
library(DiagrammeR)
library(latex2exp)
```

# Load the data

```{r source, results = "hide", fig.show='hide', cache = FALSE }
# setwd(here())
source("./GitControlled/Codes/0_0_functions.R")
# source("./GitControlled/Codes/functions_stepwise_vs_base.R")
```

```{r setup, warning=FALSE, message=FALSE, cache= TRUE}
pCorn <- price_table[2, pCorn]
pN <- price_table[2, pN]

sp_range <- 400

field <- readRDS("./Shared/Data/for_Simulations/field_padding.rds") %>%
    filter(padding==1)%>%
    dplyr::select(unique_cell_id)

####==== (1) cell-level dataset ====####
# coef_data <- readRDS("./Shared/Data/for_Simulations/cnn_coefficients_sprange_400.rds")

# x=1
# coef_data_m <- coef_data[sim == x, ]
# coef_data_t <- coef_data[sim == ifelse(x + 1 >= max(sim), 1, x + 1), ]

# sample_dt <- prepare_raw_data(1, field=field, coef_data_m=coef_data_m, coef_data_t=coef_data_t, app_error="no")

# sample_training_cell <- sample_dt$reg_raw_data

# saveRDS(sample_training_cell, "./Shared/Data/for_writing/sample_reg_raw_data.rds")

field_cell_dt <- readRDS("./Shared/Data/for_writing/sample_reg_raw_data.rds")%>%
    .[,plot_id := ceiling(subplot_id/4)]%>%
    .[padding==1,]

field_cell_sf <- left_join(field, field_cell_dt, by="unique_cell_id")%>%
    na.omit()

# ggplot(field_cell_sf)+
#     geom_sf(aes(fill=factor(padding)), size=0)

##== area of the field ==##
# st_area(st_union(field_cell_sf%>%filter(padding==1)))

# nrow(field_cell_sf)

####==== (2) plot-level dataset without padding area ====####

field_plot_sf <- field_cell_sf%>%
    group_by(plot_id, strip_id, rate)%>%
    summarise()

# nrow(field_plot_sf)
# ggplot(field_plot_sf)+geom_sf()

####==== (3) subplot-level field without padding area ====####

# field_subplot_dt <- field_cell_dt[, .(
#     sim = mean(sim),
#     yield = mean(yield),
#     opt_N = mean(opt_N),
#     rate = mean(rate),
#     aa_n = mean(aa_n), 
#     alpha = mean(alpha),
#     beta = mean(beta),
#     ymax = mean(ymax),
#     alpha1 = mean(alpha1),
#     alpha2 = mean(alpha2),
#     beta1 = mean(beta1),
#     beta2 = mean(beta2),
#     ymax1 = mean(ymax1),
#     ymax2 = mean(ymax2),
#     theta_1 = mean(theta_1),
#     theta_2 = mean(theta_2)
#   ), by = .(subplot_id, strip_id)]%>%
#   .[,unique_subplot_id:=paste0(strip_id,"_",subplot_id)]%>%
#   .[, `:=`(
#     strip_id = NULL,
#     subplot_id = NULL
#   )]

# saveRDS(field_subplot_dt, here("Shared", "Data", "for_writing", "sample_field_subplot_dt.rds"))

field_subplot_dt <- readRDS("Shared/Data/for_writing/sample_field_subplot_dt.rds")

field_subplot_sf <- field_cell_sf%>%
    group_by(subplot_id, strip_id)%>%
    summarise()%>%
    mutate(unique_subplot_id=paste0(strip_id,"_",subplot_id))%>%
    dplyr::select(unique_subplot_id)%>%
    left_join(., field_subplot_dt, by="unique_subplot_id")

# nrow(field_subplot_sf)
# ggplot(field_subplot_sf)+geom_sf()

```

# Sample CT figure

```{r}

# library(causalTree)
# library(here)
# library(rpart)
# library(rattle)

# ===================================
# Get tree 
# ===================================

# data_sf <- st_read(here("Shared/Data/snider_high_res.gpkg")) %>%
#     na.omit() %>%
#     dplyr::rename(tgtn=NPlan,tgts=SRPlan)

# data_dt <- data.table(data_sf) %>%
#     setnames(names(.),tolower(names(.)))

# tgts_ls <- data_dt[,tgts] %>% unique() %>% sort

# ctree_data <- copy(data_dt) %>%
#   .[tgts %in% tgts_ls[1:2],] %>%
#   .[,treat:=ifelse(tgts==tgts_ls[2],1,0)] %>%
#   #--- Corn Yield: 1 bu/acre ->  62.77 kg/ha---#
#   .[,yield:= yield*62.77]

# tree <- causalTree(
#   yield ~ slope+ecs,
#   data = ctree_data,
#   treatment = ctree_data$treat,
#   split.Rule = "CT",
#   cv.option = "CT",
#   split.Honest = T,
#   cv.Honest = T,
#   split.Bucket = F,
#   xval = 5,
#   cp = 0,
#   minsize = 20,
#   propensity = 0.5
# )

# opcp <- tree$cptable[,1][which.min(tree$cptable[,4])]
# opfit <- prune(tree, opcp)

# pdf(here("GitControlled/Writing/JournalSubmission/ctree_ex.pdf"))
# fancyRpartPlot(opfit,sub='')
# dev.off()

```

```{r}
variogram_tb <- data.frame(
    Parameters = c("ymax_ij", "alpha_ij", "beta_ij", "varepsilon_ij"),
    Range = c(400, 400, 400, 400),
    Mean = c(12000, -0.5, format(c(0,0))),
    Nugget = format(c(0, 0, 0, 0)),
    Sill =c(2000000, 0.02, format(1, nsmall = 1), 0.015)
    ) %>%
    flextable() %>%
    compose(i = 1, j = 1, part = "body", value = as_paragraph(as_i("ymax"), as_sup(as_i("i,j")))) %>%
    compose(i = 2, j = 1, part = "body", value = as_paragraph(as_i("\U03B1"), as_sup(as_i("i,j")))) %>%
    compose(i = 3, j = 1, part = "body", value = as_paragraph(as_i("\U03B2"), as_sup(as_i("i,j")))) %>%
    compose(i = 4, j = 1, part = "body", value = as_paragraph(as_i("\U03B5"), as_sup(as_i("i,j")))) %>%
    compose(i = 1, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
    compose(i = 1, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("6"))) %>%
    compose(i = 1, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
    compose(i = 2, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("-2"))) %>%
    compose(i = 4, j = 5, part = "body", value = as_paragraph("1.5", "\U2A2F", "10", as_sup("-2"))) %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    # --- change headers labels--- #
    set_header_labels(values = list(
        Parameters = "Parameters",
        Range = "Range (m)",
        Mean = "Mean",
        Nugget = "Nugget",
        Sill = "Sill"
    ))%>%
    #- change the borders just for consistency with other tables -#
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    footnote(
    value = as_paragraph("NOTE: About ymax, the units of of Mean, Nugget, and Sill are kg."),
     ref_symbols = NA
    ) %>%
    autofit()
```

# sample MB curve

```{r, dependson = "setup"}
# gen_yield_MB <- function(ymax, alpha, beta, N) {
#   yield <- ymax * (1 - exp(alpha + beta * N))
#   return(yield)
# }

N_rates <- field_cell_dt[,rate]%>%unique()%>%sort()

N_seq <- 0:300

        # alpha        beta     ymax
# 1: -0.6605052 -0.01572583 10886.94

dt_sample_MB_curve <- field_cell_dt[1,.(alpha,beta,ymax)]%>%
    .[rep(1:nrow(.), each = length(N_seq)), ]%>%
    .[,N:= N_seq]%>%
    .[,Yield := gen_yield_MB(ymax, alpha, beta, N)]

sample_MB_curve <- ggplot()+
    geom_point(data=dt_sample_MB_curve, 
        aes(x=N, y=Yield))+
    theme_few()+
    xlab("N(kg/ha)") + ylab("Yield(kg/ha)")

```

# Field Maps

## Division of field: plots, subplots and cells (Visualization)

```{r field-map-visualization, dependson = "setup"}

##== plot-level field without padding area ==##
# vis_field_plot <- ggplot(field_plot_sf) +
#     geom_sf(fill=NA) +
#     theme_void() +
#     labs(caption="Bottom Title") + 
#     theme(plot.title = element_text(hjust = 0.5))

####==== Example map of plot and subplots and cells ====####
##== Preparation ==##
# ex_plot <- field_plot_sf[224,]
# ex_subplots <- field_subplot_sf[ex_plot, , op = st_within]

# ex_one_subplot <- ex_subplots[1,]
# ex_cells <- field_cell_sf[ex_one_subplot, ,op=st_within]

# #-- (1) plot-level field map (base: vis_field_withoutPadding) --#
# library(latex2exp)

# plot <- 
#     ggplot()+
#         geom_sf(data=field_plot_sf)+
#         geom_sf(data=ex_plot, fill="green", size = 1)+
#         # coord_sf(expand = F) +
#         theme_void() +
#         ggtitle(TeX("$12 \\times 32$ plots")) +
#         theme(plot.title = element_text(hjust = 0.5))

# grob_plot <- ggplotGrob(plot)

# # #-- (2) create a map of subplots in a plot --##
# subplots_inPlot <- ggplot()+
#     geom_sf(data=ex_plot, fill="green")+
#     geom_sf(data = ex_subplots, fill = NA, size = 1)+
#     theme_void() +
#     ggtitle(TeX("$4 \\times 1$ subplots")) +
#     theme(plot.title = element_text(hjust = 0.5))

# grob_subplots_inPlot <- ggplotGrob(subplots_inPlot)

# # #-- (3) create a map of cells in a subplot --#
# cells_inSubplot <- 
#     ggplot()+
#     geom_sf(data=ex_one_subplot, fill="green", size = 1)+
#     geom_sf(data=ex_cells, fill=NA)+
#     theme_void() +
#     ggtitle(TeX("$6 \\times 6$ cells")) +
#     theme(
#         plot.title = element_text(hjust = 0.5, margin=margin(0,0,10,0)))

# grob_cells_inSubplot <- ggplotGrob(cells_inSubplot)

# # #-- (4)put them on the same map --#
# g_inset <- ggplot() +
#   coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE)

# map_all_together <- 
#     #/*----------------------------------*/
#     #' ##  base maps 
#     #/*----------------------------------*/
#     g_inset +
#     # --- plot level --- #
#     annotation_custom(grob_plot, 
#         xmin = 0, xmax = 0.4, ymin = 0, ymax = 1) +
#     # --- subplot level --- # 
#     annotation_custom(grob_subplots_inPlot, 
#         xmin = 0.57, xmax = 0.67, ymin = 0.3, ymax = 0.8) + #hight is 5
#     # --- cell level --- # 
#     annotation_custom(grob_cells_inSubplot, 
#         xmin = 0.8, xmax = 1, ymin = 0.5, ymax = 0.7) +
#         # xmin = 0.75, xmax = 1, ymin = 0.45, ymax = 0.7) +
#     #/*----------------------------------*/
#     #' ## add line segments
#     #/*----------------------------------*/
#     # --- plot-subplot (above) --- #
#     # geom_segment(aes(x = 0.386,  xend = 0.55, y = 0.547, yend = 0.73), 
#     geom_segment(aes(x = 0.387, xend = 0.57, y = 0.534, yend = 0.717), 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) +
#     # --- plot-subplot (below) --- #
#     # geom_segment(aes(x = 0.386, xend = 0.55, y = 0.5, yend = 0.37), 
#     geom_segment(aes(x = 0.387, xend = 0.57, y = 0.485, yend = 0.352), 
#         # data = arrowA, 
#          lineend = "round", color = "grey", 
#         size = 1.5
#         )+
#     # --- subplot-cell (above) --- #
#     # geom_segment(aes(x = 0.65, xend = 0.805, y = 0.64, yend = 0.692), 
#     geom_segment(aes(x = 0.671, xend = 0.82, y = 0.626, yend = 0.661), 
#         # data = arrowA, 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) +
#     # --- subplot-cell (below) --- #
#     geom_segment(aes(x = 0.671, xend = 0.82, y = 0.532, yend = 0.506), 
#         # data = arrowA, 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) + 
#     theme_void()

# #     theme(plot.margin=unit(c(0,0,0,0),"mm"))

# file <- here("GitControlled/Writing/Re_JournalSubmission/field_devision_v2.pdf")
# ggsave(file)
# knitr::plot_crop(file)
# dev.off()
```

# Distribution Map of Field Characteristics

## Experimental N design

```{r, dependson = "setup"}

field_Ndesign <- ggplot() +
    # geom_sf(data=field_plot_sf, color='blue', size=0.8, fill=NA) +
    geom_sf(data=field_plot_sf,
        aes(fill = factor(rate)), size = 0,
        inherit.aes = FALSE,
        # alpha=0.9
        ) +
    scale_fill_viridis_d() +
    labs(fill = "Nitrogen rate\n  (kg/ha)") +
    theme_void()
    # theme(legend.position = "bottom",
    #     legend.title = element_text(size = 0.9))

```

## Yield map

```{r, dependson = "setup"}

# yield_dist <- ggplot(field_cell_dt)+
#     geom_density(aes(x=yield))

##== cell level ==##
vis_yield_cell <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = yield), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "Yield Level\n  (kg/ha)")+
    theme_void()

##== subplot yield level ==##
vis_yield_subplot <- ggplot(field_subplot_sf) +
    geom_sf(aes(fill = yield), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "Yield Level\n  (kg/ha)")+
    theme_void()

```

## opt_N map

```{r, dependson = "setup"}
opt_N_dist <- ggplot(field_cell_dt)+
    geom_density(aes(x=opt_N))

##== cell level ==##
vis_optN_cell <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = opt_N), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "EONR (kg/ha)")+
    theme_void()

##== subplot yield level ==##
vis_optN_subplot <- ggplot(field_subplot_sf) +
    geom_sf(aes(fill = opt_N), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "Yield Level\n  (kg/ha)")+
    theme_void()
```

## Field Characteristics Map

```{r, dependson = "setup"}
####==== ymax map ====####
ymax_dist <- ggplot(field_cell_dt)+
    geom_density(aes(x=ymax))

field_ymax <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = ymax), size = 0) +
    scale_fill_viridis_c()+
    ggtitle('(1) ymax')+
    labs(fill = "kg/ha")+
    theme_void()
    # theme(legend.title = element_blank())

####==== alpha map ====####
alpha_dist <- ggplot(field_cell_dt)+
    geom_density(aes(x=alpha))

field_alpha <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = alpha), size = 0) +
    scale_fill_viridis_c() +
    ggtitle(expression("(2) " ~alpha)) +
    theme_void() +
    theme(legend.title = element_blank())

####==== beta map ====####
beta_dist <- ggplot(field_cell_dt)+
    geom_density(aes(x=beta))

field_beta <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = beta), size = 0) +
    scale_fill_viridis_c()+
    ggtitle(expression("(3) " ~beta)) +
    # labs(fill = expression(~beta))+
    theme_void() +
    theme(legend.title = element_blank())

####==== m_error map ====#### (this should be m_error*det_yield, not just m_error)
# m_error_sf <-  left_join(field, coef_data_m, by="unique_cell_id")

field_m_error <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = yield_error), size = 0) +
    scale_fill_viridis_c()+
    ggtitle(expression("(d) " ~epsilon)) +
    labs(fill = "kg/ha")+
    theme_void()
    # theme(legend.title = element_blank())

# grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

# Results

## Source Results and Preparation

```{r source-results, message=FALSE, warning=FALSE, cache= TRUE}

# ===================================
# Forest results 
# ===================================

res_forest_all_honest <- readRDS("./Shared/Results/Summary_SimRes_Honest_full_tune.rds")

# ===================================
# CNN results 
# ===================================

res_CNN_all <- readRDS("./Shared/Results/SimRes_CNN_RMSE.rds")

# /*-------------------------------------------------------*/
#' ## Organize the data
# /*-------------------------------------------------------*/ 
#--- for reporting the results of  yield prediction and EONR estimation ---#
report_res_allML <- rbind(res_forest_all_honest, res_CNN_all)%>%
    .[, Method := factor(Method, levels = c("RF", "BRF", "CNN", "CF_stepwise", "CF_base"))] %>%
    .[, Model := factor(Model, levels = c("aby", "abytt", "aabbyy", "aabbyytt"))]

report_res_subsetML <- report_res_allML %>%
    .[Method != "CF_stepwise", ]

```

# Yield prediction results

## Report R^2 of of Predicted Yield (RF, BRF, CNN)

```{r, dependson = "source-results"}

rmse_y_all <-  copy(report_res_allML)%>%
    .[Method %in% c("RF", "BRF", "CNN")]

#==== Distribution of RMSE of predicted yields ====#
plot_dis_y <- copy(rmse_y_all)%>%
    .[, Method:=factor(Method, levels = c("RF", "BRF", "CNN"))]%>%
    ggplot()+
    geom_density(aes(x=rmse_y, fill=Method), alpha=0.6)+
    facet_wrap(~Model, ncol = 1)+
    labs(x = "RMSE")+
    # labs(x = "Mean R-squared")+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

#==== Summary Table ====#
table_y_prep <- copy(rmse_y_all) %>%
    .[,.(rmse_y = mean(rmse_y)), by=.(Method, Model)] %>%
    .[,rmse_y := format(round(rmse_y,1), nsmall=1)] %>%
    dcast(Model~Method, value.var = "rmse_y")

report_table_y <- copy(table_y_prep)%>%
    .[, CF_base := "-"]%>%
    mutate(
    across(
      everything(),
      as.character
        )
    )%>%
    flextable(.) %>%
    set_header_labels(values = list(
        Model = "Model",
        RF = "RF",
        BRF = "BRF",
        CNN = "CNN",
        CF_base = "CF-base"
    ))%>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    #- change the borders just for consistency with other figures -#
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    autofit()
```

# Figure (RMSE of EONR vs RMSE of yield for RF abd BRF)

```{r}
# rmse_y_all_v2 <- rmse_y_all %>%
#     .[,`:=`(
#         rmse_y_scale = scales::rescale(rmse_y),
#         rmse_optN_scale = scales::rescale(rmse_optN)
#         ), by= .(Model, Method)]

fig_y_optN <- rmse_y_all[Method %in% c("RF", "BRF", "CNN")] %>%
    ggplot(aes(x=rmse_y, y=rmse_optN))+
    geom_point(size=0.5)+
    facet_grid(Model ~ Method) +
    # geom_smooth(method='lm', se=FALSE,  aes(colour="Regression line"), size=1) +
    # scale_colour_manual(values= "red") +
    geom_smooth(method='lm', se=FALSE,  color = "red", size=1) +
    stat_regline_equation(
        # label.x.npc = "right",
        # label.y.npc = "top", 
        label.x = 2150, label.y = 85,
        aes(label = ..rr.label..)) +
    # geom_abline(slope=1, intercept=0, color="red")+
    guides(
        fill = guide_legend(keywidth = 1, keyheight = 1),
        linetype = guide_legend(keywidth = 3, keyheight = 1),
        colour= guide_legend(keywidth = 3, keyheight = 1))+
    labs(y = " RMSE of EONR Estimation (kg/ha)")+
    labs(x = " RMSE of Yield Prediction (kg/ha)")+
    theme_few() +
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"),
        legend.title = element_blank(),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

# print(fig_y_optN, preview = "docx")
# dfabline <- data.frame(
#     x = 0:3500, y=0:3500)

# figure_te <- readRDS(here("Shared/Data/for_writing/dt_TEcomparison.rds"))%>%
#     .[, Method:= case_when(
#         Method=="RF" ~ "RF",
#         Method=="BRF" ~ "BRF",
#         Method=="CF_base" ~ "CF-base")] %>%
#     .[, Method:= factor(Method, levels = c("RF", "BRF", "CF-base"))]%>%
#     ggplot() +
#     geom_point(aes(x=true_tau_base, y=tau_base), size=0.5)+
#     # geom_abline(aes(intercept = 0, slope = 1, color = "1-to-1 line"), show.legend=TRUE) +
#     geom_line(data=dfabline, aes(x=x, y=y, color="1-to-1 line")) +
#     scale_colour_manual(labels = "1-to-1 line", values="red") +
#     facet_grid(Treatment ~ Method) +
#     guides(
#         fill = guide_legend(keywidth = 1, keyheight = 1),
#         linetype = guide_legend(keywidth = 3, keyheight = 1),
#         colour= guide_legend(keywidth = 3, keyheight = 1),
#         direction = "horizontal")+
#     labs(y = "Estimated Treatment Effect")+
#     labs(x = "True Treatment Effect")+
#     theme_few() +
#     theme(
#         strip.text.x = element_text(size=12,face="bold"),
#         strip.text.y = element_text(size=12,face="bold"),
#         legend.title = element_blank(),
#         legend.text = element_text(size=12, face="bold"),
#         legend.position = "bottom")

```

## Yield Prediction vs EONR estimation : Count Table in terms of profit loss

```{r, dependson = "source-results"}

prepare_count_tab <- rmse_y_all %>%
    dcast(sim + Model ~ Method, value.var = c("Mean", "rmse_y"))%>%
    .[,which_ML_y := case_when(
        rmse_y_BRF < rmse_y_RF & rmse_y_BRF < rmse_y_CNN ~ "BRF",
        rmse_y_RF < rmse_y_BRF & rmse_y_RF < rmse_y_CNN ~ "RF",
        rmse_y_CNN < rmse_y_BRF & rmse_y_CNN < rmse_y_RF ~ "CNN"
        )]%>%
    .[,which_ML_piLoss := case_when(
        Mean_BRF < Mean_RF & Mean_BRF < Mean_CNN ~ "BRF", 
        Mean_RF < Mean_BRF &  Mean_RF < Mean_CNN ~ "RF", 
        Mean_CNN < Mean_BRF & Mean_CNN < Mean_RF ~ "CNN"
        )] %>%
    .[,index_cnst_y_piLoss := ifelse(which_ML_piLoss==which_ML_y, 1, 0)]%>%
    .[,ML_cnst_y_piLoss:= ifelse(index_cnst_y_piLoss==1, which_ML_piLoss, NA)]

    # # RF vs BRF
    # .[,rmse_y_ratio_RF_BRF := rmse_y_RF/rmse_y_BRF]%>%
    # .[,rmse_optN_ratio_RF_BRF := rmse_optN_RF/rmse_optN_BRF]%>%
    # # CNN vs RF
    # .[,rmse_y_ratio_CNN_RF := rmse_y_CNN/rmse_y_RF]%>%
    # .[,rmse_optN_ratio_CNN_RF := rmse_optN_CNN/rmse_optN_RF]%>%
    # # CNN vs BRF 
    # .[,rmse_y_ratio_CNN_BRF := rmse_y_CNN/rmse_y_BRF]%>%
    # .[,rmse_optN_ratio_CNN_BRF := rmse_optN_CNN/rmse_optN_BRF]

# nrow(.SD[ML_cnst_y_piLoss=="BRF"])
# prepare_count_tab[ML_cnst_y_piLoss=="RF",] %>% nrow()
# prepare_count_tab[ML_cnst_y_piLoss=="BRF",] %>% nrow()
# prepare_count_tab[ML_cnst_y_piLoss=="CNN",] %>% nrow()

# Summary Table 
summary_res_CNN_RF_BRF <- prepare_count_tab%>%
    .[,.(
        count_BRF = nrow(.SD[ML_cnst_y_piLoss=="BRF",]),
        count_RF = nrow(.SD[ML_cnst_y_piLoss=="RF",]),
        count_CNN = nrow(.SD[ML_cnst_y_piLoss=="CNN",]),
        count_y_BRF = nrow(.SD[which_ML_y=="BRF",]),
        count_y_RF = nrow(.SD[which_ML_y=="RF",]),
        count_y_CNN = nrow(.SD[which_ML_y=="CNN",]),
        Total= sum(index_cnst_y_piLoss)
        ), by=Model] %>%
    .[,`:=`(
         blank1 = NA, blank2 = NA, blank3 = NA, blank4 = NA
        )] %>%
    .[,.(Model, blank1, count_y_RF, count_RF,  blank2, count_y_BRF, count_BRF, blank3, count_y_CNN, count_CNN ,blank4, Total)]

# summary_res_CNN_RF_BRF[Model=="aby", count_y_RF]

#- table creation -#
report_summary_res_CNN_RF_BRF <- copy(summary_res_CNN_RF_BRF)%>%
    flextable(.) %>%
    border_remove() %>%
    delete_part(part = "header") %>%
    add_header(
        Model="Model", 
        blank1 = "", count_y_RF = "RF", count_RF = "RF",
        blank2 = "", count_y_BRF = "BRF", count_BRF = "BRF",
        blank3 = "", count_y_CNN = "CNN", count_CNN = "CNN",
        blank4 = "", Total = "Total",
        top = TRUE) %>%
    merge_h(part = "header") %>%
    hline_bottom(j=c(3:4, 6:7, 9:10), part = "header") %>%
    add_header(
        Model="", 
        blank1 = "", count_y_RF = "#Y", count_RF = "#YP",
        blank2 = "", count_y_BRF = "#Y", count_BRF = "#YP",
        blank3 = "", count_y_CNN = "#Y", count_CNN = "#YP",
        blank4 = "", Total = "",
        top = FALSE) %>%
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    footnote(
    value = as_paragraph("NOTE: #Y indicates the number of simulation rounds in which the model provided the lowest RMSE for yield prediction. #YP indicates the number of simulation rounds in which the model provided the lower RMSE of yield prediction and the highest profit at the same time."),
     ref_symbols = NA
    ) %>%
    # set_table_properties(width = .7, layout = "autofit") %>%
    fontsize(i = NULL, j = NULL, size = 9, part = "footer") %>%
    autofit() %>%
    width(j = c(3,4,6,7,9,10,12), width=0.6)%>%
    width(j = c(2,5,8,11), width=0.1)

    # set_table_properties(width = .6, layout = "autofit")

# print(report_summary_res_CNN_RF_BRF, preview = "docx")
# in terms of RMSE of yield prediction and ", "\U1D70B\U0302", as_sub("loss"), ", respectively.

```

# EONR Estimation

```{r, dependson = "source-results"}

##== Distribution ==##
plot_dis_optN <- copy(report_res_subsetML)%>%
    .[Method %in% c("RF", "BRF", "CF_base"), ]%>%
    .[, Method := case_when(
    # Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base",
    Method == "RF" ~ "RF",
    Method == "BRF" ~ "BRF"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    # .[, Method:=factor(Method, levels = c("RF", "BRF", "CF-stepwise", "CF-base"))]%>%
    .[, Method:=factor(Method, levels = c("RF", "BRF", "CF-base"))]%>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot()+
    geom_density(aes(x=rmse_optN, fill=Method), alpha=0.7)+
    # scale_fill_viridis_d()+
    facet_wrap(~Model, ncol = 1)+
    # labs(x = expression(R^2))+
    labs(x = "RMSE (kg/ha)")+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

##== Summary Table  ==##

#- preparation -#
table_optN_prep <- copy(report_res_subsetML)%>%
    .[, .(
        rmse_optN = mean(rmse_optN),
        pi_loss = mean(Mean)
        ), by=.(Method, Model)]%>%
    .[,`:=`(
        rmse_optN = format(round(rmse_optN,1), nsmall=1),
        pi_loss = format(round(pi_loss,2), nsmall=2)
        )] %>%
    dcast(Model~Method, value.var = c("rmse_optN", "pi_loss")) %>%
    .[,`:=`(
         blank1 = NA, blank2 = NA, blank3 = NA, blank4 = NA, blank5 = NA
        )] %>%
    .[,.(Model, blank1, rmse_optN_RF, pi_loss_RF,  blank2, rmse_optN_BRF, pi_loss_BRF, blank3, rmse_optN_CNN, pi_loss_CNN ,blank4, rmse_optN_CF_base, pi_loss_CF_base)]

# table_optN_prep[Model=="aby", rmse_optN_RF]

# avg_optN <- copy(report_res_subsetML)%>%
#     .[, .(
#         Avg_rmse_optN = mean(rmse_optN),
#         Avg_pi_loss = mean(Mean)
#         ), by=Method] %>%
#     .[, `:=`(
#         Avg_rmse_optN = format(round(Avg_rmse_optN, digits=1), nsmall = 1),
#         Avg_pi_loss = format(round(Avg_pi_loss, digits=2), nsmall=2)
#         )]

#- table creation -#
report_table_optN <- copy(table_optN_prep)%>%
    mutate(
    across(
      everything(),
      as.character
        )
    )%>%
    # add_row(.after = 4) %>%
    flextable(.) %>%
    border_remove() %>%
    delete_part(part = "header") %>%
    add_header(
       Model="Model", 
        blank1 = "", rmse_optN_RF = "RF", pi_loss_RF = "RF",
        blank2 = "", rmse_optN_BRF = "BRF", pi_loss_BRF = "BRF",
        blank3 = "", rmse_optN_CNN = "CNN", pi_loss_CNN = "CNN",
        blank4 = "", rmse_optN_CF_base = "CF-base", pi_loss_CF_base = "CF-base",
        top = TRUE) %>%
    merge_h(part = "header") %>%
    hline_bottom(j=c(3:4, 6:7, 9:10, 12:13), part = "header") %>%
    add_header(
       Model="", 
        blank1 = "", rmse_optN_RF = "RMSE", pi_loss_RF = "pi_loss",
        blank2 = "", rmse_optN_BRF = "RMSE", pi_loss_BRF = "pi_loss",
        blank3 = "", rmse_optN_CNN = "RMSE", pi_loss_CNN = "pi_loss",
        blank4 = "", rmse_optN_CF_base = "RMSE", pi_loss_CF_base = "pi_loss",
        top = FALSE) %>%
    compose(i = 2, j = c(4, 7, 10, 13), part = "header", value = as_paragraph("\U1D70B\U0302", as_sub(as_i("def")))) %>%
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    # autofit() %>%
    # width(j = c(2,5,8,11), width=0.3) %>%
    # fix_border_issues() 
    footnote(
        # i = 1, j = c(4, 7, 10, 13), part = "header",
    value = as_paragraph("NOTE: \U1D70B\U0302", as_sub(as_i("def")), " indicates profit-deficit ($/ha) relative to the true maximum profit at the subplot level. The maximized profit is the profit under the true yield response functions evaluated at ", as_i("\U004E\U2071"), as_sub(as_i("opt")),"."),
     ref_symbols = NA
    ) %>%
    fontsize(i = NULL, j = NULL, size = 9, part = "footer") %>%
    autofit() %>%
    width(j = c(3,4,6,7,9,10,12,13), width=0.6)%>%
    width(j = c(2,5,8,11), width=0.1)

    # print(report_table_optN, preview = "docx")
    # autofit() 
    # width(j = c(2,5,8,11), width=0.5)
    # width(j = c(2,5,8,11), width=0.5)

# print(report_table_optN, preview = "docx")

# typology <- data.frame(
#   col_keys = names(table_optN_prep),
#   what = c(" ", " ", "RF", "RF", " ", "BRF", "BRF", " ", "CNN", "CNN", " ", "CF-base", "CF-base"),
#   measure = c(" ", " ", "RMSE", "pi_loss", " ", "RMSE", "pi_loss", " ", "RMSE", "pi_loss", " ", "RMSE", "pi_loss"),
#   stringsAsFactors = FALSE )

# copy(table_optN_prep)%>%
#     mutate(
#     across(
#       everything(),
#       as.character
#         )
#     )%>%
#     add_row(.after = 4) %>%
#     flextable(.) %>%
#     border_remove() %>%
#     add_body(
#         Model="Avg.", 
#         blank1 = NA, rmse_optN_RF = avg_optN[Method=="RF",Avg_rmse_optN], pi_loss_RF = avg_optN[Method=="RF",Avg_pi_loss],
#         blank2 = NA, rmse_optN_BRF = avg_optN[Method=="BRF",Avg_rmse_optN], pi_loss_BRF = avg_optN[Method=="BRF",Avg_pi_loss],
#         blank3 = NA, rmse_optN_CNN = avg_optN[Method=="CNN",Avg_rmse_optN], pi_loss_CNN = avg_optN[Method=="CNN",Avg_pi_loss],
#         blank4 = NA, rmse_optN_CF_base = avg_optN[Method=="CF_base",Avg_rmse_optN], pi_loss_CF_base = avg_optN[Method=="CF_base",Avg_pi_loss],
#         top = FALSE
#         )%>%
#     # delete_part(part = "header") %>%
#     set_header_df(mapping = typology, key = "col_keys") %>%
#     compose(i = 2, j = c(4, 7, 10, 13), part = "header", value = as_paragraph("\U1D70B\U0302", as_sub("loss"))) %>%
#     merge_h(part = "header") %>%
#     hline_bottom(part="all") %>%
#     hline_top(part="header") %>%
#     align(align = "center", part = "all")%>%
#     align(j=1,  align = "left", part = "all")%>%
#     autofit() %>%
#     width(j = c(2,5,8,11), width=0.5)

```

# Profit Loss

## Profit loss summary and distributions
```{r}
# piLoss_summary <- report_res_subsetML[,select(.SD, Method, Model, Min.:Max., cumulative_pi_loss)]%>%
#     .[, lapply(.SD, mean), by=.(Method, Model)] %>%
#     .[, lapply(.SD,  function(x) as.numeric(format(mean(x), digits=3))), by=.(Method, Model)] %>%
#     flextable()

piLoss_density <- 
    report_res_subsetML[Method!="CNN",] %>%
    .[,Method := case_when(
    Method == "CF_base" ~ "CF-base",
    Method == "RF" ~ "RF",
    Method == "BRF" ~ "BRF"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    .[,Method := factor(Method, levels = c("RF", "BRF", "CF-base"))] %>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot() + 
    geom_density(aes(x=Mean, fill=Method), alpha=0.5)+
    # scale_fill_viridis(discrete = TRUE, alpha=0.5) +
    # scale_fill_viridis_d()+
    # labs(x = expression(hat(pi)["loss"], " $/ha")))+
    labs(x = TeX("$\\hat{pi}_{loss}$ (\\$/ha)"))+
    facet_wrap(~Model, ncol=1) +
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")
```

# Treatment effect comparison (CF-base vs RF vs BRF)

```{r}
####=== The original code for this figure is in "1_3_CompTeEstimation.R" ===####

figure_te <- readRDS("./Shared/Data/for_writing/dt_TEcomparison.rds")%>%
    .[, Method:= case_when(
        Method=="RF" ~ "RF",
        Method=="BRF" ~ "BRF",
        Method=="CF_base" ~ "CF-base")] %>%
    .[, Method:= factor(Method, levels = c("RF", "BRF", "CF-base"))]%>%
    ggplot() +
    geom_point(aes(x=true_tau_base, y=tau_base), size=0.5)+
    geom_abline(aes(intercept = 0, slope = 1), color="red", show.legend=TRUE) +
    facet_grid(Treatment ~ Method) +
    guides(
        fill = guide_legend(keywidth = 1, keyheight = 1),
        linetype = guide_legend(keywidth = 3, keyheight = 1),
        colour= guide_legend(keywidth = 3, keyheight = 1))+
    labs(y = "Estimated Treatment Effect (kg/ha)")+
    labs(x = "True Treatment Effect (kg/ha)")+
    theme_few() +
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"),
        legend.title = element_blank(),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")
```

# Examination of Treatment Effects (CF-base vs CF-stepwise)

```{r}
# ======================================================================
# Examination of te esimation (CF-stepwise vs CF-base)
# ======================================================================

# te_res <- readRDS(here("Shared/Results/res_eval_tre_effects.rds"))%>%
#     .[,.(method, N_index, R_squared)]%>%
#     .[, N_index := case_when(
#     N_index == "1-2" ~ "(N1,N2)",
#     N_index == "2-3" ~ "(N2,N3)",
#     N_index == "3-4" ~ "(N3,N4)",
#     N_index == "4-5" ~ "(N4,N5)",
#     N_index == "1-3" ~ "(N1,N3)",
#     N_index == "1-4" ~ "(N1,N4)",
#     N_index == "1-5" ~ "(N1,N5)"
#   )]%>%
#     .[,sim:=rep(1:1000, each=8)]

##== Treatment Effects Summary Table ==##

# te_res_summary <- te_res[,.(R_squared=mean(R_squared)), by=c("method", "N_index")]%>%
#     .[, R_squared:= round(R_squared,3)]%>%
#     flextable(.)%>%
#     set_header_labels(
#     values = list(
#         method="Method",
#         N_index ="N Rates Combinations",
#         R_squared = "R-squared"
#     ))%>%
#     compose(i = 1, j = 2, part = "header", value = as_paragraph("(N", as_sup("con"), ", N", as_sup("tre"), ")"))%>%
#     compose(i = 1, j = 3, part = "header", value = as_paragraph("R",as_sup("2")))%>%
#     align(align = "center", part = "all")%>%
#     align(j=1,  align = "left", part = "all")%>%
#     # align(align = "center", part = "header")%>%
#     # align_text_col(align = "center", header = TRUE, footer = TRUE)%>%
#     autofit()

##== Graph ==##
#== tre calculation ==#

# reg_data_all <- readRDS(here("Data/for_Simulations/reg_data.rds"))
# test_agg_data_all <- readRDS(here("Data/for_Simulations/test_agg_data.rds"))

# # estimate_tre() is defined in "functions_stepwise_vs_base.R"
# sim_ex <- estimate_tre(x=32, reg_data_all=reg_data_all, test_data_all=test_agg_data_all)

# sim_ex$r2

# saveRDS(sim_ex, here("Shared", "Data","for_writing","tre_demo.rds"))

# sim_output <- readRDS(here("Shared","Data","for_writing","tre_demo.rds"))

# # r2 <-sim_output$r2
# eval <- sim_output$output

# cf_step <- eval[Method=="CF-stepwise"]%>%
#   ggplot()+
#   geom_density(aes(x=value, fill=factor(variable)),alpha=0.8)+
#   facet_wrap(~N_index, ncol = 2)+
#   theme_few()+
#   theme_few()+
#       theme(
#       legend.position="bottom",
#       plot.title = element_text(hjust = 0.5),
#       panel.background = element_rect(fill = "white"),
#         axis.ticks = element_blank(),
#         legend.title = element_blank(),
#         strip.text.x = element_text(size=12,face="bold"),
#         legend.text = element_text(size=12, face="bold"),
#         )+
#         labs(
#         x = "Treatment Effects(kg/ha)",
#         colour = "Cylinders",
#         shape = "Transmission"
#       )

# cf_base <- eval[Method=="CF-base"]%>%
#   ggplot()+
#   geom_density(aes(x=value, fill=factor(variable)),alpha=0.8)+
#   facet_wrap(~N_index, ncol = 2)+
#   theme_few()+
#   theme_few()+
#       theme(
#       legend.position="bottom",
#       plot.title = element_text(hjust = 0.5),
#       panel.background = element_rect(fill = "white"),
#         axis.ticks = element_blank(),
#         legend.title = element_blank(),
#         strip.text.x = element_text(size=12,face="bold"),
#         legend.text = element_text(size=12, face="bold"),
#         )+
#         labs(
#         x = "Treatment effects(kg/ha)",
#         colour = "Cylinders",
#         shape = "Transmission"
#       )
```

# Appendix: CF-base vs CF-stepwise

```{r, dependson = "source-results"}
# ==========================================================================
# RMSE
# ==========================================================================
##== Distribution ==##
plot_dis_CF_optN <- copy(report_res_allML)%>%
    .[Method %in% c("CF_stepwise", "CF_base"), ]%>%
    .[, Method := case_when(
    Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    .[, Method:=factor(Method, levels = c("CF-stepwise", "CF-base"))] %>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot()+
    geom_density(aes(x=rmse_optN, fill=Method), alpha=0.7)+
    # scale_fill_viridis_d()+
    facet_wrap(~Model, ncol = 1)+
    labs(x = "RMSE")+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

##== barplot ==##
CF_optN_boxplot <- copy(report_res_allML)%>%
    .[Method %in% c("CF_stepwise", "CF_base"), ]%>%
    .[, Method := case_when(
    # Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base"
    )]%>%
    .[, Method:=factor(Method, levels = c("CF-stepwise", "CF-base"))]%>%
    ggplot() + 
    geom_boxplot(aes(x=Method, y=rmse_optN, fill=Method), alpha=0.5)+
    geom_jitter(aes(x=Method, y=rmse_optN, fill=Method), size=0.3, alpha=0.9) +
    # scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    scale_fill_viridis_d()+
    labs(y = " RMSE")+
    facet_grid(~Model)+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

##== Summary Table  ==##

#- preparation -#
table_CF_optN_prep <- copy(report_res_allML)%>%
    .[, .(
        rmse_optN = mean(rmse_optN),
        pi_loss = mean(Mean)
        ), by=.(Method, Model)]%>%
    .[,`:=`(
        rmse_optN = format(round(rmse_optN,1), nsmall=1),
        pi_loss = format(round(pi_loss,2), nsmall=2)
        )] %>%
    dcast(Model~Method, value.var = c("rmse_optN", "pi_loss")) %>%
    .[,`:=`(
         blank1 = NA, blank2 = NA
        )] %>%
    .[,.(Model, blank1, rmse_optN_CF_base, pi_loss_CF_base, blank2, rmse_optN_CF_stepwise, pi_loss_CF_stepwise)]

# avg_CF_optN <- copy(report_res_allML)%>%
#     .[, .(
#         Avg_rmse_optN = mean(rmse_optN),
#         Avg_pi_loss = mean(Mean)
#         ), by=Method] %>%
#     .[, `:=`(
#         Avg_rmse_optN = format(round(Avg_rmse_optN, digits=1), nsmall = 1),
#         Avg_pi_loss = format(round(Avg_pi_loss, digits=2), nsmall=2)
#         )]

#- table creation -#
report_table_CF_optN <- copy(table_CF_optN_prep)%>%
    mutate(
    across(
      everything(),
      as.character
        )
    )%>%
    # add_row(.after = 4) %>%
    flextable(.) %>%
    border_remove() %>%
    delete_part(part = "header") %>%
    add_header(
        Model="Model", 
        blank1 = "", rmse_optN_CF_base = "CF-base", pi_loss_CF_base = "CF-base",
        blank2 = "", rmse_optN_CF_stepwise = "CF-stepwise", pi_loss_CF_stepwise = "CF-stepwise",
        top = TRUE) %>%
    merge_h(part = "header") %>%
    hline_bottom(j=c(3:4, 6:7), part = "header") %>%
    add_header(
        Model="", 
        blank1 = "", rmse_optN_CF_base = "RMSE", pi_loss_CF_base = "pi_loss",
        blank2 = "", rmse_optN_CF_stepwise = "RMSE", pi_loss_CF_stepwise = "pi_loss",
        top = FALSE) %>%
    compose(i = 2, j = c(4, 7), part = "header", value = as_paragraph("\U1D70B\U0302", as_sub("loss"))) %>%
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    # set_table_properties(width = .5, layout = "autofit")
    autofit() %>%
    width(j = c(2,5), width=0.3) %>%
    fix_border_issues()

# ==========================================================================
# profit-loss
# ==========================================================================

##== Distribution ==##
piLoss_density_CF <- copy(report_res_allML)%>%
    .[Method %in% c("CF_stepwise", "CF_base"), ]%>%
    .[, Method := case_when(
    # Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    .[, Method:=factor(Method, levels = c("CF-stepwise", "CF-base"))]%>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot() + 
    geom_density(aes(x=Mean, fill=Method), alpha=0.5)+
    # scale_fill_viridis(discrete = TRUE, alpha=0.5) +
    # scale_fill_viridis_d()+
    labs(x = expression(hat(pi)["loss"]))+
    facet_wrap(~Model, ncol=1) +
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

##== barplot ==##
CF_piLoss_boxplot <- copy(report_res_allML)%>%
    .[Method %in% c("CF_stepwise", "CF_base"), ]%>%
    .[, Method := case_when(
    # Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base"
    )]%>%
    .[, Method:=factor(Method, levels = c("CF-stepwise", "CF-base"))]%>%
    ggplot() + 
    geom_boxplot(aes(x=Method, y=Mean, fill=Method), alpha=0.5)+
    # geom_jitter(aes(x=Method, y=Mean, fill=Method), size=0.3, alpha=0.9) +
    # scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    scale_fill_viridis_d()+
    labs(y = " Mean Profit-loss")+
    facet_grid(~Model)+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

```