### 2022-02-05 Codes/parallel_reproducibility.rmd

# parallel::mclapply

+ https://www.r-bloggers.com/2020/09/future-1-19-1-making-sure-proper-random-numbers-are-produced-in-parallel-processing/

+ https://stackoverflow.com/questions/30456481/controlling-seeds-with-mclapply

+ https://stackoverflow.com/questions/40536067/how-to-adjust-future-global-maxsize-in-r

+ https://www.rdocumentation.org/packages/future/versions/1.6.1/topics/future_lapply

+ https://www.jottr.org/2017/02/19/future-rng/

```{r}
library(here)
library(mgcv)
library(ggplot2)
library(grf)
library(sf)
library(measurements)
library(data.table)
library(tidyverse)
library(future.apply)
library(parallel)
#--- source functions ---#
setwd("~/Dropbox/ResearchProject/ML_VRA")
source("./GitControlled/Codes/0_3_functions_main_sim.R")

# /*----------------------------------*/
#' ## Datasets
# /*----------------------------------*/
# === prices ===#
pCorn <- price_table[2, pCorn]
pN <- price_table[2, pN]

reg_data_all <- readRDS("./Shared/Data/for_Simulations/reg_data.rds")
test_data_all <- readRDS("./Shared/Data/for_Simulations/test_agg_data.rds")
```

## Single-core case with set.seed()
```{r}
# check whether set.seed() works for single core 
x=1

set.seed(1378)
cf_temp1 <- 
	get_opt_N(
    	reg_data = reg_data_all[sim==x&padding==1,],
    	test_data = test_data_all[sim==x&padding==1,],
    	var_ls= c("alpha","beta","ymax"),
    	rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    	Method = "CF_base"
)

set.seed(1378)
cf_temp2 <- 
	get_opt_N(
    	reg_data = reg_data_all[sim==x&padding==1,],
    	test_data = test_data_all[sim==x&padding==1,],
    	var_ls= c("alpha","beta","ymax"),
    	rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    	Method = "CF_base"
)

identical(cf_temp1, cf_temp2)
```

+ In the single-core case, set.seed() works.

## Multi-core case (mclapply) with set.seed()

```{r}
set.seed(1378)

cf_temp1 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

set.seed(1378)

cf_temp2 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

identical(cf_temp1, cf_temp2)
```

+ In the multi-core case, set.seed() does not works.

## Multi-core case (mclapply) with set.seed
+ use set.seed inside of mc.lapply

```{r}

cf_temp1 <- 
	mclapply(
		1:2, function(x) {
			set.seed(1378)

			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

cf_temp2 <- 
	mclapply(
		1:2, function(x) {
			set.seed(1378)

			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

identical(cf_temp1, cf_temp2)
```
+ This works.

# Multi-core case (mclapply) with set.seed (`RNGkind("L'Ecuyer-CMRG")`)
+ https://stackoverflow.com/questions/30456481/controlling-seeds-with-mclapply

```{r}
RNGkind("L'Ecuyer-CMRG")

set.seed(1378)

cf_temp1 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

RNGkind("L'Ecuyer-CMRG")

set.seed(1378)

cf_temp2 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2
	) %>%
	rbindlist()

identical(cf_temp1, cf_temp2)
```
+ This works.

# Multi-core case (mclapply) with set.seed (`mc.set.seed=TRUE`)

```{r}
set.seed(1378)

cf_temp1 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2,
		mc.set.seed = TRUE
	) %>%
	rbindlist()

set.seed(1378)

cf_temp2 <- 
	mclapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		mc.cores = detectCores() - 2,
		mc.set.seed = TRUE
	) %>%
	rbindlist()

identical(cf_temp1, cf_temp2)
```
+ This does not work

# Multi-core case (future_lapply) with set.seed (`future.seed = TRUE`)

```{r}
plan(multicore, workers = detectCores()-2)
options(future.globals.maxSize= 850*1024^2)

set.seed(1378)
cf_temp1 <- 
	future_lapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		future.seed = TRUE
	) %>%
	rbindlist()

set.seed(1378)
cf_temp2 <- 
	future_lapply(
		1:2, function(x) {
			get_opt_N(
    			reg_data = reg_data_all[sim==x&padding==1,],
    			test_data = test_data_all[sim==x&padding==1,],
    			var_ls= c("alpha","beta","ymax"),
    			rates_ls= reg_data_all[sim==x,]$rate%>%unique()%>%sort(),
    			Method = "CF_base"
    			)
		},
		future.seed = TRUE
	) %>%
	rbindlist()

identical(cf_temp1, cf_temp2)
```

+ This works fine. I also found that this way we can avoid using the same seed every loop unlike `mclapply()`

+ For example,

```{r}
#/*----------------------------------*/
#' ## mclapply
#/*----------------------------------*/
# RNGkind("Mersenne-Twister")

func <- function(x) runif(10)

# === 1st iteration === #
set.seed(1378)

rng1 <- unlist(
    mclapply(X = 1:10,
             FUN = function(x) runif(1),
             mc.cores = detectCores()-2,
             mc.set.seed = TRUE)
)

rng2 <- unlist(
    mclapply(X = 1:10,
             FUN = function(x) runif(1),
             mc.cores = detectCores()-2,
             mc.set.seed = TRUE)
)

identical(rng1, rng2)

x1 <- c(rng1, rng2)

# === 2nd iteration === #
set.seed(1378)

rng1 <- unlist(
    mclapply(X = 1:10,
             FUN = function(x) runif(1),
             mc.cores = detectCores()-2,
             mc.set.seed = TRUE)
)

rng2 <- unlist(
    mclapply(X = 1:10,
             FUN = function(x) runif(1),
             mc.cores = detectCores()-2,
             mc.set.seed = TRUE)
)

identical(rng1, rng2)

x2 <- c(rng1, rng2)

identical(x1, x2)

#/*----------------------------------*/
#' ## future_lapply
#/*----------------------------------*/

# === 1st iteration === #
set.seed(1378)

rng1 <- unlist(
    future_lapply(X = 1:10,
             FUN = function(x) runif(1),
             future.seed = TRUE)
)

rng2 <- unlist(
    future_lapply(X = 1:10,
             FUN = function(x) runif(1),
             future.seed = TRUE)
)

identical(rng1, rng2)

x1 <- c(rng1, rng2)

# === 2nd iteration === #
set.seed(1378)

rng1 <- unlist(
    future_lapply(X = 1:10,
             FUN = function(x) runif(1),
             future.seed = TRUE)
)

rng2 <- unlist(
    future_lapply(X = 1:10,
             FUN = function(x) runif(1),
             future.seed = TRUE)
)

identical(rng1, rng2)

x2 <- c(rng1, rng2)

identical(x1, x2)
```