### 2022-02-06 Codes/PrepareResults_1.Rmd
Correct the structure of CF_for_VRA folder

---
title: "Machine Learning Methods for Site-specific Input Management"
author: Shunkei Kakimto
output: word_document
---

```{r}
library(knitr)
# library(here)

knitr::opts_chunk$set(
    cache = FALSE,
    warning = FALSE,
    message = FALSE
    )

# opts_knit$set(root.dir=here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/ML_VRA")
```

```{r, cache = FALSE}
# === packages ===#
# --- data wrangling--- #
library(sf)
library(data.table)
library(tidyverse)

# --- figure making --- #
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(viridis)
library(ggpubr)
library(gridExtra)
library(DiagrammeR)

# --- table making --- #
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(latex2exp)
```

# Load the data

<!-- ```{r source, results = "hide", fig.show='hide', cache = FALSE }
# setwd(here())
# source("./GitControlled/Codes/0_0_functions.R")
``` -->

```{r setup, warning=FALSE, message=FALSE, cache= FALSE}
# field <- readRDS("./Shared/Data/for_Simulations/field_padding.rds") %>%
#     filter(padding==1)%>%
#     dplyr::select(unique_cell_id)

# === plot-level field data set  === #
field_plot_sf <- readRDS("./Shared/Results/for_writing/sample_field_plot_sf.rds")
# === subplot-level field data set  === #
field_subplot_sf <- readRDS("./Shared/Results/for_writing/sample_field_subplot_sf.rds")
# === cell-level field data set  === #
field_cell_sf <- readRDS("./Shared/Results/for_writing/sample_field_cell_sf.rds")
```

# Variogram Prameter Table

```{r}
variogram_tb <- data.frame(
    Parameters = c("ymax_ij", "alpha_ij", "beta_ij", "varepsilon_ij"),
    Range = c(400, 400, 400, 400),
    Mean = c(12000, -0.5, format(c(0,0))),
    Nugget = format(c(0, 0, 0, 0)),
    Sill =c(2000000, 0.02, format(1, nsmall = 1), 0.015)
    ) %>%
    flextable() %>%
    compose(i = 1, j = 1, part = "body", value = as_paragraph(as_i("ymax"), as_sup(as_i("i,j")))) %>%
    compose(i = 2, j = 1, part = "body", value = as_paragraph(as_i("\U03B1"), as_sup(as_i("i,j")))) %>%
    compose(i = 3, j = 1, part = "body", value = as_paragraph(as_i("\U03B2"), as_sup(as_i("i,j")))) %>%
    compose(i = 4, j = 1, part = "body", value = as_paragraph(as_i("\U03B5"), as_sup(as_i("i,j")))) %>%
    compose(i = 1, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
    compose(i = 1, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("6"))) %>%
    compose(i = 1, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
    compose(i = 2, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("-2"))) %>%
    compose(i = 4, j = 5, part = "body", value = as_paragraph("1.5", "\U2A2F", "10", as_sup("-2"))) %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    # --- change headers labels--- #
    set_header_labels(values = list(
        Parameters = "Parameters",
        Range = "Range (m)",
        Mean = "Mean",
        Nugget = "Nugget",
        Sill = "Sill"
    ))%>%
    #- change the borders just for consistency with other tables -#
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    footnote(
    value = as_paragraph("NOTE: About ymax, the units of of Mean, Nugget, and Sill are kg."),
     ref_symbols = NA
    ) %>%
    autofit()
```

# Field Maps

## Division of field: plots, subplots and cells (Visualization)

```{r field-map-visualization, dependson = "setup"}

##== plot-level field without padding area ==##
# vis_field_plot <- ggplot(field_plot_sf) +
#     geom_sf(fill=NA) +
#     theme_void() +
#     labs(caption="Bottom Title") + 
#     theme(plot.title = element_text(hjust = 0.5))

####==== Example map of plot and subplots and cells ====####
##== Preparation ==##
# ex_plot <- field_plot_sf[224,]
# ex_subplots <- field_subplot_sf[ex_plot, , op = st_within]

# ex_one_subplot <- ex_subplots[1,]
# ex_cells <- field_cell_sf[ex_one_subplot, ,op=st_within]

# #-- (1) plot-level field map (base: vis_field_withoutPadding) --#
# library(latex2exp)

# plot <- 
#     ggplot()+
#         geom_sf(data=field_plot_sf)+
#         geom_sf(data=ex_plot, fill="green", size = 1)+
#         # coord_sf(expand = F) +
#         theme_void() +
#         ggtitle(TeX("$12 \\times 32$ plots")) +
#         theme(plot.title = element_text(hjust = 0.5))

# grob_plot <- ggplotGrob(plot)

# # #-- (2) create a map of subplots in a plot --##
# subplots_inPlot <- ggplot()+
#     geom_sf(data=ex_plot, fill="green")+
#     geom_sf(data = ex_subplots, fill = NA, size = 1)+
#     theme_void() +
#     ggtitle(TeX("$4 \\times 1$ subplots")) +
#     theme(plot.title = element_text(hjust = 0.5))

# grob_subplots_inPlot <- ggplotGrob(subplots_inPlot)

# # #-- (3) create a map of cells in a subplot --#
# cells_inSubplot <- 
#     ggplot()+
#     geom_sf(data=ex_one_subplot, fill="green", size = 1)+
#     geom_sf(data=ex_cells, fill=NA)+
#     theme_void() +
#     ggtitle(TeX("$6 \\times 6$ cells")) +
#     theme(
#         plot.title = element_text(hjust = 0.5, margin=margin(0,0,10,0)))

# grob_cells_inSubplot <- ggplotGrob(cells_inSubplot)

# # #-- (4)put them on the same map --#
# g_inset <- ggplot() +
#   coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE)

# map_all_together <- 
#     #/*----------------------------------*/
#     #' ##  base maps 
#     #/*----------------------------------*/
#     g_inset +
#     # --- plot level --- #
#     annotation_custom(grob_plot, 
#         xmin = 0, xmax = 0.4, ymin = 0, ymax = 1) +
#     # --- subplot level --- # 
#     annotation_custom(grob_subplots_inPlot, 
#         xmin = 0.57, xmax = 0.67, ymin = 0.3, ymax = 0.8) + #hight is 5
#     # --- cell level --- # 
#     annotation_custom(grob_cells_inSubplot, 
#         xmin = 0.8, xmax = 1, ymin = 0.5, ymax = 0.7) +
#         # xmin = 0.75, xmax = 1, ymin = 0.45, ymax = 0.7) +
#     #/*----------------------------------*/
#     #' ## add line segments
#     #/*----------------------------------*/
#     # --- plot-subplot (above) --- #
#     # geom_segment(aes(x = 0.386,  xend = 0.55, y = 0.547, yend = 0.73), 
#     geom_segment(aes(x = 0.387, xend = 0.57, y = 0.534, yend = 0.717), 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) +
#     # --- plot-subplot (below) --- #
#     # geom_segment(aes(x = 0.386, xend = 0.55, y = 0.5, yend = 0.37), 
#     geom_segment(aes(x = 0.387, xend = 0.57, y = 0.485, yend = 0.352), 
#         # data = arrowA, 
#          lineend = "round", color = "grey", 
#         size = 1.5
#         )+
#     # --- subplot-cell (above) --- #
#     # geom_segment(aes(x = 0.65, xend = 0.805, y = 0.64, yend = 0.692), 
#     geom_segment(aes(x = 0.671, xend = 0.82, y = 0.626, yend = 0.661), 
#         # data = arrowA, 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) +
#     # --- subplot-cell (below) --- #
#     geom_segment(aes(x = 0.671, xend = 0.82, y = 0.532, yend = 0.506), 
#         # data = arrowA, 
#          lineend = "round", color = "grey",
#         size = 1.5
#         ) + 
#     theme_void()

# #     theme(plot.margin=unit(c(0,0,0,0),"mm"))

# file <- here("GitControlled/Writing/Re_JournalSubmission/field_devision_v2.pdf")
# ggsave(file)
# knitr::plot_crop(file)
# dev.off()
```

# Distribution Map of Field Characteristics

## Experimental N design

```{r, dependson = "setup"}
field_Ndesign <- ggplot() +
    geom_sf(data=field_plot_sf,
        aes(fill = factor(rate)), size = 0,
        inherit.aes = FALSE,
        ) +
    scale_fill_viridis_d() +
    labs(fill = "Nitrogen rate\n  (kg/ha)") +
    theme_void()
```

## Yield map

```{r, dependson = "setup"}

# === cell-level === #
vis_yield_cell <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = yield), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "Yield Level\n  (kg/ha)")+
    theme_void()

# === subplot-level === #
vis_yield_subplot <- ggplot(field_subplot_sf) +
    geom_sf(aes(fill = yield), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "Yield Level\n  (kg/ha)")+
    theme_void()
```

## opt_N map

```{r, dependson = "setup"}

# === cell-level === #
vis_optN_cell <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = opt_N), size = 0) +
    scale_fill_viridis_c()+
    labs(fill = "EONR (kg/ha)")+
    theme_void()
```

## Field Characteristics Map

```{r, dependson = "setup"}
####==== ymax map ====####
field_ymax <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = ymax), size = 0) +
    scale_fill_viridis_c()+
    ggtitle('(1) ymax')+
    labs(fill = "kg/ha")+
    theme_void()

####==== alpha map ====####
field_alpha <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = alpha), size = 0) +
    scale_fill_viridis_c() +
    ggtitle(TeX("(2)  $\\alpha$")) +
    # ggtitle(expression("(2) " ~alpha)) +
    theme_void() +
    theme(legend.title = element_blank())

####==== beta map ====####
field_beta <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = beta), size = 0) +
    scale_fill_viridis_c()+
    ggtitle(TeX("(3)  $\\beta$")) +
    theme_void() +
    theme(legend.title = element_blank())

####==== m_error map ====#### (this should be m_error*det_yield, not just m_error)
# m_error_sf <-  left_join(field, coef_data_m, by="unique_cell_id")

field_m_error <- ggplot(field_cell_sf) +
    geom_sf(aes(fill = yield_error), size = 0) +
    scale_fill_viridis_c()+
    ggtitle(TeX("(4)  $\\epsilon$")) +
    labs(fill = "kg/ha")+
    theme_void()

# grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

# Results

## Source Results and Preparation

```{r source-results, message=FALSE, warning=FALSE, cache= FALSE}

# ===================================
# Forest results 
# ===================================

res_forest_all_honest <- readRDS("./Shared/Results/Summary_SimRes_Honest_full_tune.rds")

# ===================================
# CNN results 
# ===================================

res_CNN_all <- readRDS("./Shared/Results/SimRes_CNN_RMSE.rds")

# /*-------------------------------------------------------*/
#' ## Organize the data
# /*-------------------------------------------------------*/ 
#--- for reporting the results of  yield prediction and EONR estimation ---#
report_res_allML <- rbind(res_forest_all_honest, res_CNN_all)%>%
    .[, Method := factor(Method, levels = c("RF", "BRF", "CNN", "CF_stepwise", "CF_base"))] %>%
    .[, Model := factor(Model, levels = c("aby", "abytt", "aabbyy", "aabbyytt"))]

report_res_subsetML <- report_res_allML %>%
    .[Method != "CF_stepwise", ]

```

# Yield prediction results

## Report RMSE of of Yield Prediction (RF, BRF, CNN)

```{r, dependson = "source-results"}

rmse_y_all <-  copy(report_res_allML)%>%
    .[Method %in% c("RF", "BRF", "CNN")]

#==== Distribution of RMSE of predicted yields ====#
plot_dis_y <- copy(rmse_y_all)%>%
    .[, Method:=factor(Method, levels = c("RF", "BRF", "CNN"))]%>%
    ggplot()+
    geom_density(aes(x=rmse_y, fill=Method), alpha=0.6)+
    facet_wrap(~Model, ncol = 1)+
    labs(x = "RMSE")+
    # labs(x = "Mean R-squared")+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

#==== Summary Table ====#
table_y_prep <- copy(rmse_y_all) %>%
    .[,.(rmse_y = mean(rmse_y)), by=.(Method, Model)] %>%
    .[,rmse_y := format(round(rmse_y,1), nsmall=1)] %>%
    dcast(Model~Method, value.var = "rmse_y")

report_table_y <- copy(table_y_prep)%>%
    .[, CF_base := "-"]%>%
    mutate(
    across(
      everything(),
      as.character
        )
    )%>%
    flextable(.) %>%
    set_header_labels(values = list(
        Model = "Model",
        RF = "RF",
        BRF = "BRF",
        CNN = "CNN",
        CF_base = "CF-base"
    ))%>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    #- change the borders just for consistency with other figures -#
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    autofit()
```

# Figure (RMSE of EONR vs RMSE of yield for RF abd BRF)

```{r}
fig_y_optN <- rmse_y_all[Method %in% c("RF", "BRF", "CNN")] %>%
    ggplot(aes(x=rmse_y, y=rmse_optN))+
    geom_point(size=0.5)+
    facet_grid(Model ~ Method) +
    # geom_smooth(method='lm', se=FALSE,  aes(colour="Regression line"), size=1) +
    # scale_colour_manual(values= "red") +
    geom_smooth(method='lm', se=FALSE,  color = "red", size=1) +
    stat_regline_equation(
        # label.x.npc = "right",
        # label.y.npc = "top", 
        label.x = 2150, label.y = 85,
        aes(label = ..rr.label..)) +
    # geom_abline(slope=1, intercept=0, color="red")+
    guides(
        fill = guide_legend(keywidth = 1, keyheight = 1),
        linetype = guide_legend(keywidth = 3, keyheight = 1),
        colour= guide_legend(keywidth = 3, keyheight = 1))+
    labs(y = " RMSE of EONR Estimation (kg/ha)")+
    labs(x = " RMSE of Yield Prediction (kg/ha)")+
    theme_few() +
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"),
        legend.title = element_blank(),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

# print(fig_y_optN, preview = "docx")
# dfabline <- data.frame(x = 0:3500, y=0:3500)

```

## Yield Prediction vs EONR estimation : Count Table in terms of profit loss

```{r, dependson = "source-results"}

prepare_count_tab <- rmse_y_all %>%
    dcast(sim + Model ~ Method, value.var = c("Mean", "rmse_y"))%>%
    .[,which_ML_y := case_when(
        rmse_y_BRF < rmse_y_RF & rmse_y_BRF < rmse_y_CNN ~ "BRF",
        rmse_y_RF < rmse_y_BRF & rmse_y_RF < rmse_y_CNN ~ "RF",
        rmse_y_CNN < rmse_y_BRF & rmse_y_CNN < rmse_y_RF ~ "CNN"
        )]%>%
    .[,which_ML_piLoss := case_when(
        Mean_BRF < Mean_RF & Mean_BRF < Mean_CNN ~ "BRF", 
        Mean_RF < Mean_BRF &  Mean_RF < Mean_CNN ~ "RF", 
        Mean_CNN < Mean_BRF & Mean_CNN < Mean_RF ~ "CNN"
        )] %>%
    .[,index_cnst_y_piLoss := ifelse(which_ML_piLoss==which_ML_y, 1, 0)]%>%
    .[,ML_cnst_y_piLoss:= ifelse(index_cnst_y_piLoss==1, which_ML_piLoss, NA)]

# === Summary Table === #
summary_res_CNN_RF_BRF <- prepare_count_tab%>%
    .[,.(
        count_BRF = nrow(.SD[ML_cnst_y_piLoss=="BRF",]),
        count_RF = nrow(.SD[ML_cnst_y_piLoss=="RF",]),
        count_CNN = nrow(.SD[ML_cnst_y_piLoss=="CNN",]),
        count_y_BRF = nrow(.SD[which_ML_y=="BRF",]),
        count_y_RF = nrow(.SD[which_ML_y=="RF",]),
        count_y_CNN = nrow(.SD[which_ML_y=="CNN",]),
        Total= sum(index_cnst_y_piLoss)
        ), by=Model] %>%
    .[,`:=`(
         blank1 = NA, blank2 = NA, blank3 = NA, blank4 = NA
        )] %>%
    .[,.(Model, blank1, count_y_RF, count_RF,  blank2, count_y_BRF, count_BRF, blank3, count_y_CNN, count_CNN ,blank4, Total)]

# === Table creation === #
report_summary_res_CNN_RF_BRF <- copy(summary_res_CNN_RF_BRF)%>%
    flextable(.) %>%
    border_remove() %>%
    delete_part(part = "header") %>%
    add_header(
        Model="Model", 
        blank1 = "", count_y_RF = "RF", count_RF = "RF",
        blank2 = "", count_y_BRF = "BRF", count_BRF = "BRF",
        blank3 = "", count_y_CNN = "CNN", count_CNN = "CNN",
        blank4 = "", Total = "Total",
        top = TRUE) %>%
    merge_h(part = "header") %>%
    hline_bottom(j=c(3:4, 6:7, 9:10), part = "header") %>%
    add_header(
        Model="", 
        blank1 = "", count_y_RF = "#Y", count_RF = "#YP",
        blank2 = "", count_y_BRF = "#Y", count_BRF = "#YP",
        blank3 = "", count_y_CNN = "#Y", count_CNN = "#YP",
        blank4 = "", Total = "",
        top = FALSE) %>%
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    footnote(
    value = as_paragraph("NOTE: #Y indicates the number of simulation rounds in which the model provided the lowest RMSE for yield prediction. #YP indicates the number of simulation rounds in which the model provided the lower RMSE of yield prediction and the highest profit at the same time."),
     ref_symbols = NA
    ) %>%
    # set_table_properties(width = .7, layout = "autofit") %>%
    fontsize(i = NULL, j = NULL, size = 9, part = "footer") %>%
    autofit() %>%
    width(j = c(3,4,6,7,9,10,12), width=0.6)%>%
    width(j = c(2,5,8,11), width=0.1)
```

# EONR Estimation

```{r, dependson = "source-results"}

# === Distribution of RMSE of EONR estimates === #

plot_dis_optN <- copy(report_res_subsetML)%>%
    .[Method %in% c("RF", "BRF", "CF_base"), ]%>%
    .[, Method := case_when(
    # Method == "CF_stepwise" ~ "CF-stepwise",
    Method == "CF_base" ~ "CF-base",
    Method == "RF" ~ "RF",
    Method == "BRF" ~ "BRF"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    # .[, Method:=factor(Method, levels = c("RF", "BRF", "CF-stepwise", "CF-base"))]%>%
    .[, Method:=factor(Method, levels = c("RF", "BRF", "CF-base"))]%>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot()+
    geom_density(aes(x=rmse_optN, fill=Method), alpha=0.7)+
    # scale_fill_viridis_d()+
    facet_wrap(~Model, ncol = 1)+
    # labs(x = expression(R^2))+
    labs(x = "RMSE (kg/ha)")+
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")

# === Summary Table === #

# --- prepation --- #
table_optN_prep <- copy(report_res_subsetML)%>%
    .[, .(
        rmse_optN = mean(rmse_optN),
        pi_loss = mean(Mean)
        ), by=.(Method, Model)]%>%
    .[,`:=`(
        rmse_optN = format(round(rmse_optN,1), nsmall=1),
        pi_loss = format(round(pi_loss,2), nsmall=2)
        )] %>%
    dcast(Model~Method, value.var = c("rmse_optN", "pi_loss")) %>%
    .[,`:=`(
         blank1 = NA, blank2 = NA, blank3 = NA, blank4 = NA, blank5 = NA
        )] %>%
    .[,.(Model, blank1, rmse_optN_RF, pi_loss_RF,  blank2, rmse_optN_BRF, pi_loss_BRF, blank3, rmse_optN_CNN, pi_loss_CNN ,blank4, rmse_optN_CF_base, pi_loss_CF_base)]

#- table creation -#
report_table_optN <- copy(table_optN_prep)%>%
    mutate(
    across(
      everything(),
      as.character
        )
    )%>%
    # add_row(.after = 4) %>%
    flextable(.) %>%
    border_remove() %>%
    delete_part(part = "header") %>%
    add_header(
       Model="Model", 
        blank1 = "", rmse_optN_RF = "RF", pi_loss_RF = "RF",
        blank2 = "", rmse_optN_BRF = "BRF", pi_loss_BRF = "BRF",
        blank3 = "", rmse_optN_CNN = "CNN", pi_loss_CNN = "CNN",
        blank4 = "", rmse_optN_CF_base = "CF-base", pi_loss_CF_base = "CF-base",
        top = TRUE) %>%
    merge_h(part = "header") %>%
    hline_bottom(j=c(3:4, 6:7, 9:10, 12:13), part = "header") %>%
    add_header(
       Model="", 
        blank1 = "", rmse_optN_RF = "RMSE", pi_loss_RF = "pi_loss",
        blank2 = "", rmse_optN_BRF = "RMSE", pi_loss_BRF = "pi_loss",
        blank3 = "", rmse_optN_CNN = "RMSE", pi_loss_CNN = "pi_loss",
        blank4 = "", rmse_optN_CF_base = "RMSE", pi_loss_CF_base = "pi_loss",
        top = FALSE) %>%
    compose(i = 2, j = c(4, 7, 10, 13), part = "header", value = as_paragraph("\U1D70B\U0302", as_sub(as_i("def")))) %>%
    hline_bottom(part="all") %>%
    hline_top(part="header") %>%
    align(align = "center", part = "all")%>%
    align(j=1,  align = "left", part = "all")%>%
    # autofit() %>%
    # width(j = c(2,5,8,11), width=0.3) %>%
    # fix_border_issues() 
    footnote(
        # i = 1, j = c(4, 7, 10, 13), part = "header",
    value = as_paragraph("NOTE: \U1D70B\U0302", as_sub(as_i("def")), " indicates profit-deficit ($/ha) relative to the true maximum profit at the subplot level. The maximized profit is the profit under the true yield response functions evaluated at ", as_i("\U004E\U2071"), as_sub(as_i("opt")),"."),
     ref_symbols = NA
    ) %>%
    fontsize(i = NULL, j = NULL, size = 9, part = "footer") %>%
    autofit() %>%
    width(j = c(3,4,6,7,9,10,12,13), width=0.6)%>%
    width(j = c(2,5,8,11), width=0.1) 
```

# Profit Loss

## Profit loss summary and distributions
```{r}
piLoss_density <- 
    report_res_subsetML[Method!="CNN",] %>%
    .[,Method := case_when(
    Method == "CF_base" ~ "CF-base",
    Method == "RF" ~ "RF",
    Method == "BRF" ~ "BRF"
    )]%>%
    .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
    )]%>%
    .[,Method := factor(Method, levels = c("RF", "BRF", "CF-base"))] %>%
    .[, Model:=factor(Model, 
        levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt"))] %>%
    ggplot() + 
    geom_density(aes(x=Mean, fill=Method), alpha=0.5)+
    # scale_fill_viridis(discrete = TRUE, alpha=0.5) +
    # scale_fill_viridis_d()+
    # labs(x = expression(hat(pi)["loss"], " $/ha")))+
    labs(x = TeX("$\\hat{pi}_{loss}$ (\\$/ha)"))+
    facet_wrap(~Model, ncol=1) +
    theme_few()+
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")
```

# Treatment effect comparison (CF-base vs RF vs BRF)

```{r}
####=== The original code for this figure is in "1_3_CompTeEstimation.R" ===####

figure_te <- readRDS("./Shared/Results/for_writing/dt_TEcomparison.rds")%>%
    .[, Method:= case_when(
        Method=="RF" ~ "RF",
        Method=="BRF" ~ "BRF",
        Method=="CF_base" ~ "CF-base")] %>%
    .[, Method:= factor(Method, levels = c("RF", "BRF", "CF-base"))]%>%
    ggplot() +
    geom_point(aes(x=true_tau_base, y=tau_base), size=0.5)+
    geom_abline(aes(intercept = 0, slope = 1), color="red", show.legend=TRUE) +
    facet_grid(Treatment ~ Method) +
    guides(
        fill = guide_legend(keywidth = 1, keyheight = 1),
        linetype = guide_legend(keywidth = 3, keyheight = 1),
        colour= guide_legend(keywidth = 3, keyheight = 1))+
    labs(y = "Estimated Treatment Effect (kg/ha)")+
    labs(x = "True Treatment Effect (kg/ha)")+
    theme_few() +
    theme(
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"),
        legend.title = element_blank(),
        legend.text = element_text(size=12, face="bold"),
        legend.position = "bottom")
```