### 2022-02-06 Presentation/DIFM/presentation_x.rmd

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: "Taro Mieno, University of Nebraska Lincoln"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
---
class: middle

```{r, child = 'setup.Rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
#--- load packages ---#
suppressMessages(library(data.table))
suppressMessages(library(tidyverse))
suppressMessages(library(DT))
suppressMessages(library(sf))
suppressMessages(library(here))
suppressMessages(library(flextable))
suppressMessages(library(printr))
suppressMessages(library(tidycensus))
suppressMessages(library(tmap))
suppressMessages(library(RefManageR))

here::i_am("GitControlled/Presentation/ML_VRA.bib")

```

```{r tmap-layout, cache = FALSE}
tm_layout_to_add <-
  tm_layout(
    legend.outside = "TRUE",
    frame = FALSE,
    legend.title.size = 2,
    legend.text.size = 1.5
  )
```

```{r setup, include=FALSE, cache = FALSE}
options(htmltools.dir.version = FALSE)
BibOptions(
  check.entries = FALSE,
  bib.style = "authoryear",
  style = "markdown",
  dashed = TRUE,
  max.names = 1,
  longnamesfirst = FALSE
)
file.name <- here("GitControlled/Presentation", "ML_VRA.bib")
bib <- ReadBib(file.name)
```

```{r figure_setup, cache = F}
theme_update(
  axis.title.x =
    element_text(
      size = 9, angle = 0, hjust = .5, vjust = -0.3, face = "plain"
    ),
  axis.title.y =
    element_text(
      size = 9, angle = 90, hjust = .5, vjust = .9, face = "plain"
    ),
  axis.text.x =
    element_text(
      size = 7, angle = 0, hjust = .5, vjust = 1.5, face = "plain"
    ),
  axis.text.y =
    element_text(
      size = 7, angle = 0, hjust = 1, vjust = 0, face = "plain"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 7, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.title =
    element_text(
      size = 7, angle = 0, hjust = 0, vjust = 0, face = "plain"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 7),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 9),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r analyses, include = FALSE, cache = TRUE}
# /*=================================================*/
#' # Preparation
# /*=================================================*/
# This is where you prepare data sets and results

# /*----------------------------------*/
#' ## Create figures and tablesb
# /*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl(here("GitControlled/Codes/PrepareResults.Rmd"), output = here("GitControlled/Codes/PrepareResults.R"))

# --- Read the simulation results and prepare figures and tables ---#
source(here("GitControlled/Codes/PrepareResults.R"))
```

# Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

+ Shunkei Kakimoto, University of Nebraska Lincoln
+ Taro Mieno, University of Nebraska Lincoln 
+ Takashi Tanaka, Gifu University, Japan
+ David S. Bullock, University of Illinois

---
class: middle

# Site-specific Economically Optimal Input Rates

+ On-farm Precision Experimentation (OFPE)
  - more spatially granular data of yield, inputs of interest (e.g., fertilizer, seed)
  - other field/soil characteristics
+ Steps:
  1. Run input experiments
  2. Collect data
  3. <span style = "color: red;"> Estimate yield response function based on the data </span>  
  4. Estimate the economically optimal site-specific input rates

---
class: middle

# Yield Response and EOIR Estimation

+ Application of various machine learning methods
  - Random Forest `r Citep(bib, c("krause2020random", "MARQUESRAMOS2020105791"))`
  - Convolutional Neural Network `r Citep(bib, "barbosa2020modeling")`
+ Many studies
  + stop at predicting yields without identifying EOIR
  + use information that is not available to farmers at the time of their management decision making

---
class: middle

# Yield and EOIR prediction

Consider this stylized conceptual model:

$$Yield = f(I, c, x) + g(c, x)$$

+ <span style = "color: red;"> $f(I, c, x)$ </span>: $0.4$ at maximum
+ $g(c, x)$: $0.6$ at maximum

```{r }
data.table(
  model = rep(c("model 1", "model 2", "model 3"), each = 2),
  component = rep(c("f", "g"), 3),
  explained = c(0.3, 0.5, 0.3, 0.6, 0.4, 0.4),
  `R-squared` = c(0.8, 0.8, 0.9, 0.9, 0.8, 0.8)
) %>%
  flextable() %>%
  color(i = c(3, 4), j = 1, color = "blue") %>%
  color(i = c(5, 6), j = 1, color = "red") %>%
  merge_v(j = c(1, 4)) %>%
  hline(i = c(2, 4, 6)) %>%
  fix_border_issues() %>%
  align(j = c(2, 3, 4), align = "center") %>%
  autofit()

```

---
class: middle

.content-box-red[**Key**]: You want to estimate accurately the part that matters for EOIR suggestion, not overall.

---
class: middle

# Predicting Yield <span style = "color: red;"> Level </span> is Sufficient, but not Necessary

+ Consider a very simple profit maximization problem

$$\max_{I} \quad pf(I, \mathbf{c}) - wI$$

+ $p$: crop price
+ $f(\cdot)$: yield function
+ $I$: input
+ $c$: collection of other variables
+ $w$: input price

--

.content-box-red[**First order condition**]:

$$\frac{\partial f(I^{*}(p,w,\mathbf{c}), \mathbf{c})}{\partial I} \equiv \frac{w}{p}$$

What you need is the impact of input on yield (difference in yield moving from one rate to another)

---
class: middle

# Causal Forest and other traditional ML methods

+ Traditional ML methods (e.g., Random Forest, Boosted Regression Forest, CNN) 
  - is designed to estimate the <span style = "color: red;"> level </span> of the dependent variable (yield) well
  - once the yield model is trained, you can estimate the impact of a treatment (just plug in two different rates and see the difference in their predicted yield values)

+ Causal Forest (a causal machine learning method)
  - is designed to estimate heterogeneous treatment effects directly without predicting yield levels

---
class: middle

# Causal Forest: Illustration (Corn, Seed Rate Experiment)

.left4[
<br>
+ A causal forest is an ensemble of causal trees
+ Treatment effect is calculated at each leaf

<br>

.content-box-blue[Experiment:]

+ **Control**: $27,000$ seeds/acre
+ **Treated**: $31,000$ seeds/acre

.content-box-blue[Characteristics:]
+ Slope (`slope`)
+ Electrical Conductivity (`ecs`) 
]

.right6[
```{r c-tree, echo = F, out.width = "100%"}
include_graphics(here("GitControlled/Presentation", "ctree.png"))
```
]

---
class: middle

# Research Question

+ Does Causal Forest (CF) outperform traditional ML approaches in estimating site-specific EOIR and result in higher economic profits?

+ Can you trust yield prediction accuracy as the criteria to select a model when your ultimate goal is to find the model that produces the highest economic profit?

---
class: middle

# Method

+ Monte Carlo simulation
  - hypothetical field, on which OFPEs are run
  - site-specific yield response functions are known (to researchers)
  - site-specific economically optimal input rates are known for a given combination of crop and input prices

<span style = "color: red;"> Important Note </span>: Using real-world data, you are never ever able to test which model performs the best in terms of the economic performance of site-specific input recommendations because you do not observe true yield response functions

---
class: middle

# Yield Response Functions

Yield response function is assumed to follow Mitscherlich-Baule functional form:

$$f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon$$

+ $c$: a collection of site-specific characteristics
+ $\varepsilon$: error

--

Researchers observe and use yield, N, and $c$ to train models.

---
class: middle

# Data Generation: Nitrogen Rate (Trial Design)
.left4[
<br>
<br>
Each plot is assigned a single N rate according to the fixed trial design
]

.right6[
```{r fig-nExp, out.width="100%", cache = TRUE}
field_Ndesign
```
]

---
class: middle

# Data Generation: Characteristics and Error

.left4[
<br>
<br>
Characteristics $(c)$ and error term are generated so that they are spatially correlated 
]

.right6[
```{r fig-parameters, out.width="100%", cache = TRUE}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol = 2, nrow = 2)
```
]

---
class: middle

# Data Generation: Yield

.left4[
<br>
<br>
Yield is generated based on N, $c$, and error term
]

.right6[
```{r field-yield-cell, out.width="100%", cache = TRUE}
vis_yield_cell
```
]

---
class: middle

# Modeling Scenarios

Researchers observe all the variables (characteristics) that govern site-specific yield response functions, but do not know the functional form.

$$f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) N)) + \varepsilon$$

+ **aby**: $\alpha$, $\beta$, ymax 
+ **abytt**: $\alpha$, $\beta$, ymax, $\theta_1$, $\theta_2$ $(\theta$s are irrelevant variables)
+ **aabbyy**: 
  - $\alpha_1$, $\alpha_2$ $(\alpha = \alpha_1 + \alpha_2)$ 
  - $\beta_1$, $\beta_2$ $(\beta = \beta_1 + \beta_2)$ 
  - $ymax_1$, $ymax_2$ $(ymax = ymax_1 + ymax_2)$
+ **aabbyytt**: $\alpha_1$, $\alpha_2$, $\beta_1$, $\beta_2$, $ymax_1$, $ymax_2$, $\theta_1$, $\theta_2$

---
class: middle

# Estimation Methods

+ Random Forest
+ Boosted Regression Forest
+ Convolutional Neural Network (follows the architecture of `r Citet(bib, "barbosa2020modeling")`)
+ Causal Forest (CF)

---
class: middle

# Results: Economic Performance

```{r  echo = F, out.width = "80%"}
include_graphics(here("GitControlled/Presentation", "results_table.png"))
```

---
class: middle

# Treatment Effect Estimation Accuracy: subplot-level

```{r plot-tre, fig.dim = c(6, 4), cache = TRUE}
figure_te
```

---
class: middle

# Yield Prediction vs Profitability Performance

```{r  echo = F, out.width = "80%"}
include_graphics(here("GitControlled/Presentation", "yield_profit.png"))
```

---
class: middle

# Conclusions

## Economic Performance
+ Random Forest should not be used. It is simply worse than BRF and CF and vulnerable to irrelevant variables.
+ CF outperforms RF, BRF, CNN
+ **Caveats**:
  * CNN used a particular architecture suggested by `r Citep(bib, "barbosa2020modeling")`, and other architectures may perform better
  * If the experimental settings differ (yield response functions, trial design, etx), the results can of course be different (I already found a case where BRF performs better than CF just yesterday)

--

## Yield Prediction vs EONR Prediction
+ Yield prediction accuracy has very little to do with EONR prediction accuracy
+ Picking a model based on yield prediction accuracy does not necessarily lead you to a model that is good at predicting EONR

---
class: middle

# Extensions

+ Collaborate across the teams to see which model works the best under what condition?
+ Can we identify factors that lets us know what approach to use with some confidence?