### 2022-02-06 Presentation/OFE_Presentation/presen_slides.rmd

---
title: Machine Learning Methods for Site-specific Input Management
author:
  - Shunkei Kakimoto$^1$, Taro Mieno$^1$ and Takashi S. T. Tanaka$^2$
institute: $^1$Agricultural Economics, University of Nebraska Lincoln  $^2$Applied Biological Sciences, Gifu University
date: "September 1st 2021"
fontsize: 12pt
mathspec: true # for metropolis theme
aspectratio: 169 # 16:9
#classoption: handout
output:
 beamer_presentation:
   theme: "Madrid" 
   # https://deic-web.uab.cat/~iblanes/beamer_gallery/index_by_theme.html
   #theme: "metropolis" # special
   colortheme: "whale"
   # https://deic-web.uab.cat/~iblanes/beamer_gallery/index_by_color.html
   fonttheme: "professionalfonts"
   # https://deic-web.uab.cat/~iblanes/beamer_gallery/index_by_font.html
   latex_engine: xelatex # for metropolis theme
   toc: false
   number_sections: false
   slide_level: 1
   keep_tex: true
    # includes:
    #   in_header: preamble.tex
# compact-title: ture # to remove markdown generated title frame
header-includes:
  - \usepackage{fancyvrb}
  - \usepackage{graphicx}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  # - \setbeamertemplate{page number in head/foot}[]{}
  # - \AtBeginSection{}
bibliography: MSthesis.bib
---

```{r setup, include=FALSE}
library(here)
library(knitr)

# opts_knit$set(root.dir = here())

opts_chunk$set(
  fig.align = "center", 
  fig.retina = 5,
  warning = F, 
  message = F,
  cache = FALSE, # <-- T
  cache.lazy = FALSE,
  echo = F
  )
```

```{r, cache = TRUE}
# === packages ===#
library(sf)
library(data.table)
library(grf)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(tidyverse)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(modelsummary)
library(raster)
library(bookdown)
```

# Background
+ The application of Machine Learning(ML) methods for site-specific economically optimal input rates (EOIR) (e.g., seed, fertilizer) have been getting more attention in recent years
\newline
  + @barbosa2020modeling applied Convolutional Neural Network
  + @krause2020random used Random Forest-based approaches
  + @Gardner21, @Wang2021, @coulibali2020, etc.

---

# Research Gap

+ The conventional ML methods focus on predicting yield well rather than causal identification of input on yield
+ The past studies have used predictive ability of yield for validity of their models.
\newline

## note
+ EOIR estimation should be based on the change in yields associated with the change in input levels *ceteris paribus*
+ Having good yield prediction capability does not necessarily mean it is also capable of estimating EOIR well
<!-- + Because causal identification between yield and input is necessary for estimation of EONR, and such causal inference and having good prediction is different two distinct object -->

---

# New Trend of Causal Machine Learning Application 
+ Causal Machine Learning (CML) methods:
  + Unlike the conventional prediction-oriented ML methods, CML focuses on identifying causal impacts of an event (in our context, an increase or decrease in input rate for example)
  \newline
+ **Causal Forest (CF) ** [@Wager2018a; @athey2016recursive]:
  + CF estimates heterogeneous causal impacts of a treatment (a change in the input level) based on observed characteristics (e.g., organic matter)
<!-- + Hence, CF seems particularly suitable for estimating site-specific yield response functions and then EOIR.  -->

---

# Research Questions

##
+ In terms of estimating EOIR, how do CF-based methods (CF-stepwise and CF-base) compare to other prediction-oriented ML methods: Random Forest (RF), Boosted Random Forest (BRF), and Convolutional Neural Network (CNN)?
\newline
+ Is the predictive ability of yield a good indicator of the performance of EOIR estimation?

---

# Methods
+ Conduct one thousand rounds of Monte Carlo simulations under four different production scenarios
+ Compare the performance in estimating economically optimal nitrogen rates (EONR) of CF-based methods (CF-base and CF-stepwise) to other methods: RF, BRF, and CNN
+ For RF, BRF, and CNN, contrast their EONR performances against their yield prediction performances
<!-- +  Evaluation metric: R-squared of estimated EONR and R-squared of predicted yield -->

---

# Key Results: EONR estimation

```{r res-N, out.width = "70%"}
# knit_print(table_optN_subplot_wide)
knitr::include_graphics("table_optN.png")
```

---

# Key Results: EONR estimation and Yield Prediction

```{r res-yield, out.width = "70%"}
# knit_print(table_optN_subplot_wide)
knitr::include_graphics("table_y.png")
```
---

# Key Findings
## 
+ The proposed CF-base method is capable of estimating site-specific EONR more accurately than other prediction-oriented ML methods 
\newline
+ The ability of predicting yield does not necessarily translate to good performance in estimating EONR.

---

# References {.allowframebreaks}