### 2022-02-06 Writing/CEA_submission/Highlights.rmd
restructuring of the folders and files

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto, Taro Mieno, Takashi S. T. Tanaka, and David S. Bullock
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: no
    global_numbering: true
    reference_docx: word_template.docx
    # reference_docx: word_template_format.docx
    # pandoc_args: ["-Fpandoc-crossref"]
bibliography: ML_VRA.bib
csl: computers-and-electronics-in-agriculture.csl
eqnPrefix: 
      - "Eq"
---

# Highlights:

+ Causal Forest is a novel machine learning method focusing on the identification of heterogeneous causal effects of treatment unlike the conventional machine learning methods.

+ Causal Forest outperforms Random Forest, Boosted Random Forest, and Convolutional Neural Network in estimating site-specific EONR.

+ The model's yield prediction accuracy provides little if any information about its EONR prediction accuracy.