### 2022-03-15 Writing/CEA_2nd_round/replies_rev_1.rmd
CEA 3rd round

---
title: "Reply to Reviewer #1"
output:
  bookdown::word_document2:
    number_sections: no
    reference_docx: ../word_template_format.docx
---

Thank you so much for your comments. By addressing them, we believe the manuscript has improved significantly. Below you will find your comments in bold and our replies in regular font.

<br>

**1. The first point of the highlight describes the advantages of the CF model over other machine learning methods. The advantages of the CF model are not suitable for the highlight of this study. Therefore, I suggest that the authors revise the first point of the highlight.**

<span style="color:blue">  
We agree with your point, and we have dropped the discussion of the advantages of the CF model from our highlights.
</span>

<br>

**2. In the second section (lines 76-184), the authors describe the fundamental information necessary to identify EOIR in general and discuss the causal forest approach in detail. The structure of the manuscript is redundant. Some of the descriptions belong to the methods section. Therefore, I suggest that the authors adjust the content of the section. Move some content to the supplement material and others to the methods section.**
<span style="color:blue">

<br>

We do see a redundancy in our explanation of how CF can be used to identify site-specific input recommendations. It was explained twice, once each in subsections 2.2 and 3.2.2. We have deletedthe explanation in 2.2 and also changed the heading in 2.1 to "EOIR Estimation and Treatment Effects."

We have moved an example of a causal tree (previously in seciton 2.3), but have kept the explanation of how CF works. The core objective of the paper is to test the CF against existing ML methods in predicting EOIR and economic performance of resulting site-specific input use recommendations.

ML simulation is the core method we use to pursue this core objective. For this reason the Materials and Methods section features no explanation of how CF works. We also provide little explanation of RF, BRF, and CNN methods, which are well-established and familiar to many readers of this journal. However, since this article is the first to apply CF in the field of agronomy, let alone site-specific input management, we believe that it is worthwhile to present in detail how it works and the conditions necessary for it to work well. We believe that if we placed such a description of how CF works inside the Materials and Method section, it will unnecessarily distract the reader by breaking the flow of the explanation of the MC simulation procedures. We feel that the narrative will flow better if it first informs readers about the potential advantages of CF over existing prediction-oriented ML methods before moving into a description of the MC simulations in the Materials and Method section. Therefore, we have kept in the current manuscript the discussion located in subsection 2.3 and 2.4 of the original manuscript.
</span>

<br>

**3. The description of materials and methods section is redundant, please move some unimportant details description to supplementary materials.**

<span style="color:blue">  
In rereponse to this comments, we made several changes.

First, in the new manuscript we consolidated the original manuscript's 3.1.2 and 3.1.3 into a single subsubsection called "Cell-specific Yield Functions and True Economically Optimal N Rates." In doing so, we moved to the supplementary appendix (Appendix B) the detailed explanation of the conceptual framework behind the yield response functions based on cell-level characteristics, $c_i$. In addition, we moved to Appendix C our detailed discussion of parameters and error term generation, which included a number of tables and figures (Table C.1, Figure C.1).

Second, we consolidated subsection 3.1.4 and 3.1.5 of the original manuscript into a single subsubsection called "Trial Design and Yield Data Generation for Modeling." In doing so, we removed some text and combined the original manuscript's figures 5 and 7 in the single figure 2 in the current manuscript.

Third, we moved to Appendix C.2 the previous manuscript’s detailed description of parameter splitting.

We believe that the detailed presentation of data generation is crucial in our reporting of MC procedures, as it would be in the report of any empirical study using real experiment data provides. In both cases, results are dependent procedures. We believe further slimming of the description would hinder anyone attempting to reproduce our results. That said, we have worked hard to remove redundancies in this section of the manuscript. 
</span>

<br>

**4. What is the unit of the topographical slope？ (line 131)**

<span style="color:blue">  
It is degree. The unit is now provided in the description of the figure in Appendix A. Please note that the purpose of the example is simply to illustrate how CF works.
</span>

<br>

**5. There are 3 numbers in each small frame of Figure 1, please describe the meaning of each number in detail in the figure caption. (line 137)**

<span style="color:blue">  
Explanations for those numbers are added as a footnote to the figure. The figure is now in an appendix and called Figure A.1. 
</span>

<br>

**6. Some formulas in the manuscript are not numbered (lines 116, 120, 268, 291, 371, 389).**

<span style="color:blue">  
All the formulas are now numbered.
</span>

<br>

**7. What are the exact number and proportion of test set training sets? (line 192)**

<span style="color:blue">  
We apologize for not making this cleaner in the original manuscript. The training data set and test data set were completely separate in our analysis. For a given simulation round, the training data set from another round acted as a "test" dataset. These data sets were created independently and with different underlying field parameters $\alpha$, $\beta$, and $ymax$. Training data sets, however, were split into several parts for cross-validation in determining a round's optimal values of its hyperparameters. In addition, we added the following paragraph section 3.3.3 to make this point clear.

"For evaluation of yield and EONR predictions, a data set from a different simulation round was used as the test data set. For example, for simulation number 1, the training data set from simulation number 2 was used as the test data set. This process works because the training data sets were generated by the same data generating process, yet have different underlying field parameters ($\alpha$, $\beta$, $ymax$). Further, all training data sets were generated independently, so for any particular simulation round, the data set from any other round could serve as a valid test data set. This practice made the creation of one thousand additional simulations fields unnecessary."
</span>

<br>

**8. I can't open the code link (https://github.com/tmieno2/ML_VRA.git, lines 199 and 249) with my browsers (chrome, edge, and Firefox). Please upload the code to the public research data repositories (Zenodo, Mendeley Data, figshare, Harvard Dataverse, etc.), not GitHub. Please generate a separate DOI for your code and cite it.**

<span style="color:blue">  
We have followed the Editor's suggestion by sharing the simulation codes and the data on Mendeley Data (https://doi.org/10.17632/y7cdhzww6w.1) to increase the accessibility. The name of the data set is "Codes and Data Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation", and two folders are included in this data set: "Codes" and  "Data". The "Code" folder contains all the simulation codes, and the "Data" folder contains a boundary polygon data of a field which is necessary to create field data for this study. Please read README.md for the steps to reproduce. In addition, we also created a public github repository named VRA_with_CF (https://github.com/Shunkei3/VRA_with_CF.git).
</span>

<br>

**9. Please convert the length units of the manuscript to international length units. (line 207)**

<span style="color:blue">  
We modified the length units (line 163) 
</span>

<br>

**10. Please add the number of rows and columns to the plots, subplots, and cells schematics in Figure 2. (line 209)**

<span style="color:blue">  
We modified the figure. Please see Figure 1. 
</span>

<br>

**11. The titles of sections 3.2.1 and 3.2.2 do not provide useful information to the readers, please revise the titles. (lines 342 and 373)**

<span style="color:blue">  
We changed the title of 3.2 to "Training Models and Estimating Site-specific EONRs", and the title of 3.2.1 and 3.2.2 to "Training and Estimation of Site-specific EONRs by RF, BRF, and CNN" and "Training and Estimation of Site-specific EONRs by CF".
</span>

<br>

**12. The titles of 4.1 and 4.2 are confusing, please revise. (lines 410 and 457)**

<span style="color:blue">  
To better reflect the contents of 4.1, we changed its title to "Comparison of EONR Estimation Accuracy." We changed the title of 4.2 to "The Relationship between Yield and EONR Prediction Performances." Both new headings provide more accurate descriptions of their subsection's contents.
</span>

<br>

**13. Why there are no CNN results in Figures 8,9, and 10? (lines 439, 441, and 453)**

<span style="color:blue">

As explained in section 4.1, CNN failed to predict site-specific EONR because the estimated yield response function were always linear (Table 1). Therefore, we thought that visualizing the CNN results would not provide much insight. However, we have decided to include CNN results in Figures 8, 9, and 10 of the previous manuscript. which are Figures 3, 4, and 5 of the revised manuscript.

</span>

<br>

**14. I suggest that the authors add to the supplementary material a table of the training accuracy of the different models.**

<span style="color:blue">  
We have newly added the results of training accuracy to Appendix D. 
</span>

<br>

**15. Please add a separate section for research gaps or uncertainties. Are there any prerequisite assumptions for using the CF model?**

<span style="color:blue">  
We have added a section called “Limitations and Future Studies” before the Conclusion section of the revised manuscript. One of the important conditions for CF to provide consistent estimation of treatment effects is that the treatment (N rates here) is uncorrelated with unobservable characteristics. Other assumptions are relatively minor. Note that, this is not just for CF, but also for RF, BRF, and CNN. Whenever the variable of interest (N rates here) is correlated with the unobservables (error term), their estimation of its impact on the dependent variable is biased in general. Fortunately, in OFPE settings we can randomize or geographically pattern input rates to make them uncorrelated with unobserved variables. Therefore, the violation of this assumption is not as much of a concern as in a setting where observational data is used. We discuss these in the “Trial Design and Yield Data Generation for Modeling” section. 
</span>

<br>

**16. Please revise the conclusion. Most of the content is discussing the uncertainty of the study. Please move the statement about uncertainty to a separate uncertainty section.**

<span style="color:blue">  
As mentioned earlier in our reply to your comment 14, in the revised manuscript we have moved the descriptions of the limitations (uncertainty) of the study to a separate section. 
</span>