### 2022-03-15 Writing/CEA_2nd_round/replies_rev_2.rmd

---
title: "Reply to Reviewer #2"
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: no
    global_numbering: true
    reference_docx: ../word_template_format.docx
---

Thank you so much for your comments. By addressing them, we believe the manuscript has improved significantly. Below you will find your comments in bold and our replies in regular font.

<br>

**1. All tables and figs: No capitalization of the titles, other than first letter.**

<span style="color:blue"> 
We modified all the titles of tables and figures according to your suggestions.
</span>

**2. Fig, 1. Indicate units for slope and EC in title.**

<span style="color:blue"> 
The units of slope and EC are added in the description, which can be found in Appendix A.
</span>

**3. Table 1 and Fig. 3: show units where applicable.**
<span style="color:blue">

Units are added to the table and figure where applicable. They are now in Appendix C.
</span>

**4. p. 21 l. 334: For example, instead of using , directly, some some topographical land features and soil properties (remove 1 some)**

<span style="color:blue"> 
The extra "some" was removed from the sentence (see p15 l.250).
</span>

**5. p. 24 l. 403: The accuracy of EONR estimation was judged based on the RMSE (/h)of EONR estimation**

<span style="color:blue"> 
The typo was corrected (see p19 l.321). 
</span>

**6. p. 26 l. 424: To examine whether this problem is a product solely of the specific choice**

<span style="color:blue">  
The grammatical error was corrected (see p20 l.341). 
</span>

**7. P. 29 l 449: Points are clustered much more tightly about the 1-to-1 line much for CF-base than RF and BRF. (remove much)**

<span style="color:blue">  
We removed "much" from the sentence (see p24 l.367).
</span>

**8. Table 3. Provide units.**

<span style="color:blue"> 
The unit (kg/ha) was provided in the table (it is now Table 2). Similarly, units are provided to the other result tables.
</span>