### 2022-03-15 Writing/CEA_3rd_round/replies_rev_1.rmd

---
title: "Reply to Editor"
output:
  bookdown::word_document2:
    number_sections: no
    reference_docx: ../word_template_format.docx
---

Thank you so much for your comments. By addressing them, we believe the manuscript has improved significantly. Below you will find your comments in bold and our replies in regular font.

<br>

**1. The CNN result seems to be rather constant than linear. Please include a chart showing the CNN performance per training epoch to allow the reader to assess how good the CNN model converged.**

<span style="color:blue">

</span>

<br>

**2. Line 70. Why can't EOIRs be estimated by the intervention studies mentioned in 28ff?**

<span style="color:blue">

</span>

<br>

**3. Line 194 simualtions**

<span style="color:blue">

</span>

<br>

**4. Line 194 How exactly were the simulation parameters (table c.1, figure c.1) inspired by the field study (line 161ff) ?**

<span style="color:blue">

</span>

<br>

**5. Line 321 It seems RSME was never defined.**

<span style="color:blue">

</span>

<br>

**6. In principle, your ML models have to learn to replicate equation 3, right? You have to explain better why RSME for yield and EONR can be uncorrelated when they are directly related through Eq. 3 and 5. There must be an effect of yield prediction error on EONR. By definition, a 100% perfect yield prediction should reduce the EONR error to the effect of epsilon in Eq. 3. (As a side note, in this case the correlation between the RSMEs would also be zero.) A larger RSME for yield prediction would simultaneously increase the RSME of EONR until the EONR virtually becomes a random variable. Your comparison in Figure 6 may not be insightful in this regard. It might be confounded by effects from different prediction performance for different simulation parameters i.e. the prediction performance might be different depending on where you are in the hyperbola from Eq.3 . If the prediction performance is not uniform or the prediction error is of a similar size as the achievable yield increase, then it is natural that the RSMEs of EONR and yield prediction are not correlated. Figure C1.4 indicates that the effect size of the random error epsilon alone is already in the range of the achievable yield increase. I recommend to add a simulation of different ML performances and different epsilon values to test their effects.**

<span style="color:blue">

</span>

<br>

**7. Please insert a plot for true yield vs yield prediction to enable the reader to estimate the achieved prediction performance.**

<span style="color:blue">

</span>

<br>

**8. Please show to which degree N affected yield and how the relation between yield estimation error and the effect of N on yield increase was. I would like to see whether the models were able to reproduce the effect of N on yield at all.**

<span style="color:blue">

</span>

<br>

**9. Please extend the introduction by restating numbers for the state of the art of yield and EONR prediction performance from previous studies (as far as such numbers are available) to prevent that all of your readers have to consult the cited references.**

<span style="color:blue">

</span>

<br>

**10. Please discuss your results with regard to the state of the art (again as specific as possible) and discuss whether your yield prediction performance and your simulation approach (e.g. epsilon effect size) may have affected the results for EONR prediction.**

<span style="color:blue">

</span>

<br>