### 2022-06-09 Writing/CEA_final_submission/Highlights.rmd

---
title: "Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation"
author: Shunkei Kakimoto, Taro Mieno, Takashi S. T. Tanaka, and David S. Bullock
output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: no
    global_numbering: true
    reference_docx: ../word_template.docx
---

# Highlights:

+ Causal Forest is a novel machine learning method focusing on the identification of heterogeneous causal effects of treatment unlike the conventional machine learning methods.

+ Causal Forest outperforms Random Forest, Boosted Random Forest, and Convolutional Neural Network in estimating site-specific EONR.

+ The model's yield prediction accuracy provides little if any information about its EONR prediction accuracy.